---
title: "NMAT Foreign Examinees Dashboard, 2006-2018"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    code_folding: hide
    self_contained: true
fig_width: 11
fig_height: 7
---

```{r css, echo=FALSE, results='asis'}
cat('
<style>
.metric-row { display:flex; flex-wrap:wrap; gap:12px; margin:16px 0; }
.metric-card {
  background:#fff; border-left:4px solid #3c8dbc; border-radius:4px;
  padding:12px 18px; min-width:170px; box-shadow:0 1px 3px rgba(0,0,0,.12);
}
.metric-value { font-size:1.6em; font-weight:700; color:#2c3e50; }
.metric-label { font-size:.82em; color:#7f8c8d; margin-top:2px; }
.callout-info {
  background:#e8f4f8; border-left:4px solid #3498db; padding:10px 16px;
  margin:12px 0; border-radius:2px; font-size:.92em;
}
.callout-warning {
  background:#fff8e1; border-left:4px solid #f39c12; padding:10px 16px;
  margin:12px 0; border-radius:2px; font-size:.92em;
}
.section-hdr {
  background:#f0f4ff; border-left:3px solid #1a237e; padding:8px 14px;
  margin:18px 0 10px 0; border-radius:2px; font-size:.95em;
  font-weight:600; color:#1a237e;
}
hr.break { border:none; border-top:2px solid #e0e0e0; margin:22px 0; }
.small-note { color:#666; font-size:12px; margin:-6px 0 12px 0; }
</style>
')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(plotly)
  library(DT)
  library(scales)
  library(RColorBrewer)
  library(htmltools)
  library(knitr)
})

options(scipen = 999)

BINORDER <- paste0("B", 1:10)
PALBIN <- c(
  B1 = "#8B0000", B2 = "#B22222", B3 = "#D9534F", B4 = "#F0AD4E", B5 = "#FFD166",
  B6 = "#A0D468", B7 = "#66C2A5", B8 = "#41B6C4", B9 = "#2C7FB8", B10 = "#253494"
)

SUBTESTSTD <- c(
  Verbal = "NMS_VCss",
  Inductive = "NMS_IRss",
  Quantitative = "NMS_Qss",
  Perceptual = "NMS_PAss",
  Biology = "NMS_BIOss",
  Physics = "NMS_PHYss",
  Social = "NMS_SSCss",
  Chemistry = "NMS_CHEMss"
)

`%||%` <- function(a, b) {
  if (!is.null(a) && length(a) > 0) a else b
}

squish_chr <- function(x) {
  x <- as.character(x)
  x <- gsub("\\s+", " ", trimws(x))
  x[x %in% c("", "NA", "N/A", "NULL", "Na")] <- NA_character_
  x
}

title_case_clean <- function(x) {
  x <- squish_chr(x)
  out <- ifelse(is.na(x), NA_character_, tools::toTitleCase(tolower(x)))
  out
}

safe_mean <- function(x) {
  if (all(is.na(x))) return(NA_real_)
  round(mean(x, na.rm = TRUE), 2)
}

safe_median <- function(x) {
  if (all(is.na(x))) return(NA_real_)
  round(median(x, na.rm = TRUE), 2)
}

safe_q <- function(x, p) {
  if (all(is.na(x))) return(NA_real_)
  round(as.numeric(stats::quantile(x, p, na.rm = TRUE, names = FALSE)), 2)
}

pal_n <- function(n, palette = "Set2") {
  if (n <= 0) return(character(0))
  base_n <- max(3, min(8, n))
  colorRampPalette(brewer.pal(base_n, palette))(n)
}

sanitize_for_dt <- function(df) {
  df <- as.data.frame(df, stringsAsFactors = FALSE)
  for (col in names(df)) {
    x <- df[[col]]
    if (is.factor(x)) df[[col]] <- as.character(x)
    if (is.list(x)) {
      df[[col]] <- vapply(
        x,
        function(v) {
          if (is.null(v) || length(v) == 0) return(NA_character_)
          paste(as.character(v), collapse = ", ")
        },
        character(1)
      )
    }
  }
  df
}

dtexport <- function(data, caption = NULL) {
  data <- sanitize_for_dt(data)
  datatable(
    data,
    caption = if (!is.null(caption)) tags$caption(
      style = "font-weight:bold;font-size:13px;text-align:left;",
      caption
    ),
    extensions = c("Buttons", "Scroller"),
    options = list(
      dom = "Blfrtip",
      buttons = list(
        list(extend = "copy", text = "Copy"),
        list(extend = "csv", text = "CSV", filename = gsub("[^A-Za-z0-9]+", "_", caption %||% "NMAT_export")),
        list(extend = "excel", text = "Excel", filename = gsub("[^A-Za-z0-9]+", "_", caption %||% "NMAT_export"), title = caption %||% "NMAT Export"),
        list(extend = "print", text = "Print")
      ),
      scrollX = TRUE,
      scrollY = "320px",
      scroller = TRUE,
      pageLength = 25,
      lengthMenu = list(c(25, 50, 100, -1), c(25, 50, 100, "All"))
    ),
    rownames = FALSE,
    class = "display compact stripe"
  )
}

metric_card <- function(value, label, color = "#3c8dbc") {
  div(
    class = "metric-card",
    style = paste0("border-left-color:", color, ";"),
    div(class = "metric-value", value),
    div(class = "metric-label", label)
  )
}

metric_row <- function(...) {
  div(class = "metric-row", ...)
}

find_foreigners_csv <- function() {
  candidates <- c(
    "REAL_FOREIGNERS.csv",
    "REAL_FOREIGNERS-2.csv",
    "dataset/REAL_FOREIGNERS.csv",
    "dataset/REAL_FOREIGNERS-2.csv",
    "../REAL_FOREIGNERS.csv",
    "../REAL_FOREIGNERS-2.csv",
    "../dataset/REAL_FOREIGNERS.csv",
    "../dataset/REAL_FOREIGNERS-2.csv",
    "../../REAL_FOREIGNERS.csv",
    "../../REAL_FOREIGNERS-2.csv",
    "../../dataset/REAL_FOREIGNERS.csv",
    "../../dataset/REAL_FOREIGNERS-2.csv"
  )

  for (p in candidates) {
    if (file.exists(p)) return(p)
  }

  recursive_hits <- c(
    list.files(".", pattern = "^REAL_FOREIGNERS.*\\.csv$", recursive = TRUE, full.names = TRUE, ignore.case = TRUE),
    list.files("..", pattern = "^REAL_FOREIGNERS.*\\.csv$", recursive = TRUE, full.names = TRUE, ignore.case = TRUE),
    list.files("../..", pattern = "^REAL_FOREIGNERS.*\\.csv$", recursive = TRUE, full.names = TRUE, ignore.case = TRUE)
  )
  recursive_hits <- recursive_hits[file.exists(recursive_hits)]

  if (length(recursive_hits) > 0) return(recursive_hits)[1]

  stop("Cannot find REAL_FOREIGNERS.csv / REAL_FOREIGNERS-2.csv. Place the CSV in the NMAT_Shiny folder, project root, or dataset/ folder.")
}

normalize_center <- function(x) {
  y <- squish_chr(x)
  z <- tolower(y)

  z[z %in% c("los angeles  (us)", "los angeles (us)", "los angeles(us)", "los angeles  (u.s.)", "los angeles (u.s.)")] <- "Los Angeles (US)"
  z[z %in% c("las vegas  (us)", "las vegas (us)", "las vegas(us)", "las vegas  (u.s.)", "las vegas (u.s.)")] <- "Las Vegas (US)"
  z[z %in% c("cagayan de oro", "cagayan de oro city")] <- "Cagayan De Oro"

  out <- ifelse(
    is.na(z),
    NA_character_,
    ifelse(grepl("\\(us\\)$", z, fixed = FALSE),
           gsub("\\(us\\)$", "(US)", tools::toTitleCase(z)),
           tools::toTitleCase(z))
  )
  out <- gsub("\\(Us\\)", "(US)", out)
  out
}

normalize_nationality <- function(x) {
  y <- squish_chr(x)
  z <- tolower(y)
  z <- gsub("[[:punct:]]", " ", z)
  z <- gsub("\\s+", " ", trimws(z))
  out <- ifelse(is.na(z), NA_character_, NA_character_)

  out[z %in% c("india", "indian")] <- "India"
  out[z %in% c("nepal", "nepali")] <- "Nepal"
  out[z %in% c("thailand", "thai")] <- "Thailand"
  out[z %in% c("nigeria", "nigerian")] <- "Nigeria"
  out[z %in% c("american", "united states", "usa", "u s a", "u s", "us")] <- "United States"

  fill_idx <- is.na(out) & !is.na(z)
  out[fill_idx] <- tools::toTitleCase(z[fill_idx])

  out <- gsub("\\bUsa\\b", "USA", out)
  out <- gsub("^Uk$", "UK", out)
  out
}

normalize_school_type <- function(x) {
  y <- title_case_clean(x)
  y[y %in% c("Private School", "Private University")] <- "Private"
  y[y %in% c("Public School", "Public University", "State")] <- "Public"
  y[is.na(y)] <- "Not Specified"
  y
}

normalize_course_group <- function(x) {
  y <- title_case_clean(x)
  y[is.na(y)] <- "Not Specified"
  y
}

make_percentile_bin <- function(x) {
  cut(
    x,
    breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 101),
    labels = BINORDER,
    right = FALSE,
    include.lowest = TRUE
  )
}

pctcross <- function(df, rowcol, colcol, colorder = NULL) {
  sub <- df %>%
    filter(!is.na(.data[[rowcol]]), !is.na(.data[[colcol]])) %>%
    count(.data[[rowcol]], .data[[colcol]], name = "n") %>%
    pivot_wider(names_from = all_of(colcol), values_from = n, values_fill = 0L)

  if (!is.null(colorder)) {
    missing_cols <- setdiff(colorder, names(sub))
    for (m in missing_cols) sub[[m]] <- 0L
    keep <- c(rowcol, intersect(colorder, names(sub)))
    sub <- sub[, keep, drop = FALSE]
  }

  counts <- sub
  if (nrow(counts) == 0) return(list(counts = counts, pct = counts))

  numeric_part <- counts[, setdiff(names(counts), rowcol), drop = FALSE]
  row_totals <- rowSums(numeric_part, na.rm = TRUE)
  pct_part <- round(sweep(numeric_part, 1, pmax(row_totals, 1), "/") * 100, 2)
  pct <- cbind(counts[rowcol], pct_part)

  list(counts = counts, pct = pct)
}

kwtable <- function(df, groupcol, colmap) {
  bind_rows(lapply(names(colmap), function(col) {
    if (!(col %in% names(df)) || !(groupcol %in% names(df))) return(NULL)
    sub <- df %>% filter(!is.na(.data[[col]]), !is.na(.data[[groupcol]]))
    groups <- split(sub[[col]], sub[[groupcol]])
    groups <- Filter(function(g) length(g) > 1, groups)
    if (length(groups) < 2) return(NULL)

    kt <- tryCatch(kruskal.test(groups), error = function(e) NULL)
    if (is.null(kt)) return(NULL)

    k <- length(groups)
    n <- nrow(sub)
    eta2 <- max(0, as.numeric(kt$statistic - (k - 1)) / max(n - k, 1))

    tibble(
      Score = colmap[[col]],
      H = round(as.numeric(kt$statistic), 3),
      p_value = ifelse(kt$p.value < 0.001, "<0.001", sprintf("%.4f", kt$p.value)),
      eta_squared = round(eta2, 4)
    )
  }))
}

subtestmeans <- function(df, groupcol) {
  cols <- SUBTESTSTD
  mask <- unname(cols) %in% names(df)
  colvals <- unname(cols)[mask]
  colnms <- names(cols)[mask]
  if (length(colvals) == 0 || !(groupcol %in% names(df))) return(NULL)

  out <- df %>%
    filter(!is.na(.data[[groupcol]])) %>%
    group_by(.data[[groupcol]]) %>%
    summarise(across(all_of(colvals), ~ round(mean(.x, na.rm = TRUE), 2)), .groups = "drop")

  for (i in seq_along(colvals)) {
    j <- which(names(out) == colvals[i])
    if (length(j) > 0) names(out)[j] <- colnms[i]
  }

  names(out)[1] <- groupcol[1]
  out
}

subtestheatmap <- function(tbl, title = "Subtest Heatmap") {
  if (is.null(tbl) || nrow(tbl) == 0) return(plotly_empty())
  gcol <- names(tbl)[1]
  mat <- as.matrix(tbl[, -1, drop = FALSE])
  rownames(mat) <- as.character(tbl[[gcol]])

  plot_ly(
    z = mat,
    x = colnames(mat),
    y = rownames(mat),
    type = "heatmap",
    colorscale = "RdBu",
    reversescale = TRUE,
    hovertemplate = "Group: %{y}<br>Subtest: %{x}<br>Mean: %{z}<extra></extra>"
  ) %>%
    layout(
      title = title,
      xaxis = list(title = "Subtest"),
      yaxis = list(title = "")
    )
}

radarplot_tbl <- function(tbl, title = "Radar Plot") {
  if (is.null(tbl) || nrow(tbl) == 0) return(plotly_empty())
  gcol <- names(tbl)[1]
  cats <- names(tbl)[-1]

  fig <- plot_ly(type = "scatterpolar", fill = "toself")
  for (i in seq_len(nrow(tbl))) {
    vals <- as.numeric(unlist(tbl[i, -1, drop = TRUE], use.names = FALSE))
    fig <- fig %>%
      add_trace(
        r = c(vals, vals[1]),
        theta = c(cats, cats[1]),
        name = as.character(tbl[[gcol]][i])
      )
  }

  fig %>%
    layout(
      title = title,
      polar = list(radialaxis = list(visible = TRUE, range = c(350, 700))),
      showlegend = TRUE
    )
}

yearly_summary_foreign <- function(df) {
  df %>%
    filter(!is.na(Year)) %>%
    group_by(Year) %>%
    summarise(
      n = n(),
      permedian = safe_median(NMSPERnum),
      perq25 = safe_q(NMSPERnum, 0.25),
      perq75 = safe_q(NMSPERnum, 0.75),
      gpsmedian = safe_median(NMSGPS),
      gpsq25 = safe_q(NMSGPS, 0.25),
      gpsq75 = safe_q(NMSGPS, 0.75),
      aptmedian = safe_median(NMSAPT),
      samedian = safe_median(NMSSA),
      subcomp_median = safe_median(SubtestCompositeStd),
      .groups = "drop"
    )
}

load_and_prep <- function() {
  fp <- find_foreigners_csv()
  df <- read.csv(fp, stringsAsFactors = FALSE, check.names = FALSE)

  numeric_cols <- intersect(
    c(
      "Year", "NMA_TestDate", "NMA_YearGrad",
      "NMS_VCss", "NMS_IRss", "NMS_Qss", "NMS_PAss",
      "NMS_BIOss", "NMS_PHYss", "NMS_SSCss", "NMS_CHEMss",
      "NMS_APT", "NMS_SA", "NMS_GPS", "NMS_PER"
    ),
    names(df)
  )

  for (col in numeric_cols) {
    df[[col]] <- suppressWarnings(as.numeric(df[[col]]))
  }

  if ("Year" %in% names(df)) df$Year <- as.integer(df$Year)
  if ("NMA_AppNo" %in% names(df)) df$NMA_AppNo <- squish_chr(df$NMA_AppNo)
  if ("NMA_Name" %in% names(df)) df$NMA_Name <- squish_chr(df$NMA_Name)

  df$NMSPERnum <- if ("NMS_PER" %in% names(df)) suppressWarnings(as.numeric(df$NMS_PER)) else NA_real_
  df$NMSGPS <- if ("NMS_GPS" %in% names(df)) suppressWarnings(as.numeric(df$NMS_GPS)) else NA_real_
  df$NMSAPT <- if ("NMS_APT" %in% names(df)) suppressWarnings(as.numeric(df$NMS_APT)) else NA_real_
  df$NMSSA <- if ("NMS_SA" %in% names(df)) suppressWarnings(as.numeric(df$NMS_SA)) else NA_real_

  df$SEXCLEAN <- case_when(
    "NMA_Sex" %in% names(df) & as.character(df$NMA_Sex) == "1" ~ "Male",
    "NMA_Sex" %in% names(df) & as.character(df$NMA_Sex) == "2" ~ "Female",
    TRUE ~ NA_character_
  )

  df$NATIONALITY_RAW <- if ("NAC_NATIONALITY" %in% names(df)) squish_chr(df$NAC_NATIONALITY) else NA_character_
  df$NATIONALITY_CLEAN <- normalize_nationality(df$NATIONALITY_RAW)

  df$CENTER_RAW <- if ("NMC_Center" %in% names(df)) squish_chr(df$NMC_Center) else NA_character_
  df$CENTER_CLEAN <- normalize_center(df$CENTER_RAW)

  df$COLLEGE_CLEAN <- if ("NMA_College" %in% names(df)) title_case_clean(df$NMA_College) else NA_character_
  df$SCHOOL_TYPE <- if ("School Type_rec2_FINAL" %in% names(df)) normalize_school_type(df$`School Type_rec2_FINAL`) else "Not Specified"
  df$COURSE_CLASS_CLEAN <- if ("Course Classification" %in% names(df)) normalize_course_group(df$`Course Classification`) else "Not Specified"
  df$COURSE_RECODE_CLEAN <- if ("Course_recode" %in% names(df)) normalize_course_group(df$Course_recode) else "Not Specified"
  df$LOCAL_PROVINCE <- if ("NMAT Province local address" %in% names(df)) title_case_clean(df$`NMAT Province local address`) else NA_character_
  df$PERM_REGION <- if ("NMAT Region permanent address" %in% names(df)) title_case_clean(df$`NMAT Region permanent address`) else NA_character_

  df$IS_OVERSEAS_PERM <- !is.na(df$PERM_REGION) & df$PERM_REGION == "Overseas"
  df$IS_OVERSEAS_LOCAL <- !is.na(df$LOCAL_PROVINCE) & df$LOCAL_PROVINCE == "Overseas"

  df$PercentileBin <- make_percentile_bin(df$NMSPERnum)
  df$PercentileBin <- factor(df$PercentileBin, levels = BINORDER, ordered = TRUE)

  std_cols_present <- intersect(unname(SUBTESTSTD), names(df))
  if (length(std_cols_present) > 0) {
    mat <- as.matrix(df[, std_cols_present, drop = FALSE])
    comp <- rowMeans(mat, na.rm = TRUE)
    comp[apply(is.na(mat), 1, all)] <- NA_real_
    df$SubtestCompositeStd <- round(comp, 2)
    df$HasAllStdSubtests <- rowSums(is.na(mat)) == 0
  } else {
    df$SubtestCompositeStd <- NA_real_
    df$HasAllStdSubtests <- FALSE
  }

  df
}
```

```{r data_prep}
DFRAW <- load_and_prep()
DFTREND <- DFRAW %>% filter(!is.na(Year), Year >= 2006, Year <= 2018)

YEAR_SUM <- yearly_summary_foreign(DFTREND)

MISSING_TBL <- tibble(
  column = names(DFRAW),
  missing_n = vapply(DFRAW, function(x) sum(is.na(x)), numeric(1)),
  missing_pct = round(vapply(DFRAW, function(x) mean(is.na(x)) * 100, numeric(1)), 2)
) %>% arrange(desc(missing_pct), desc(missing_n))

QUALITY_TBL <- tibble(
  Check = c(
    "Rows in foreigner dataset",
    "Columns in foreigner dataset",
    "Unique application numbers",
    "Duplicate application numbers",
    "Rows with percentile",
    "Rows with GPS",
    "Rows with all 8 standardized subtests",
    "Rows with nationality",
    "Rows with testing center",
    "Rows with college"
  ),
  Count = c(
    nrow(DFTREND),
    ncol(DFTREND),
    DFTREND %>% filter(!is.na(NMA_AppNo)) %>% summarise(n = n_distinct(NMA_AppNo)) %>% pull(n),
    DFTREND %>% filter(!is.na(NMA_AppNo)) %>% count(NMA_AppNo) %>% filter(n > 1) %>% nrow(),
    sum(!is.na(DFTREND$NMSPERnum)),
    sum(!is.na(DFTREND$NMSGPS)),
    sum(DFTREND$HasAllStdSubtests, na.rm = TRUE),
    sum(!is.na(DFTREND$NATIONALITY_CLEAN)),
    sum(!is.na(DFTREND$CENTER_CLEAN)),
    sum(!is.na(DFTREND$COLLEGE_CLEAN))
  )
)

HARMONIZATION_TBL <- tibble(
  Dimension = c("Nationality", "Testing center"),
  Raw_distinct = c(
    n_distinct(na.omit(DFTREND$NATIONALITY_RAW)),
    n_distinct(na.omit(DFTREND$CENTER_RAW))
  ),
  Clean_distinct = c(
    n_distinct(na.omit(DFTREND$NATIONALITY_CLEAN)),
    n_distinct(na.omit(DFTREND$CENTER_CLEAN))
  ),
  Reduction = Raw_distinct - Clean_distinct
)

NAT_RAW_CT <- DFTREND %>%
  filter(!is.na(NATIONALITY_RAW)) %>%
  count(NATIONALITY_RAW, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

NAT_CT <- DFTREND %>%
  filter(!is.na(NATIONALITY_CLEAN)) %>%
  count(NATIONALITY_CLEAN, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

CENTER_CT <- DFTREND %>%
  filter(!is.na(CENTER_CLEAN)) %>%
  count(CENTER_CLEAN, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

SCHOOL_CT <- DFTREND %>%
  count(SCHOOL_TYPE, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

COLLEGE_CT <- DFTREND %>%
  filter(!is.na(COLLEGE_CLEAN)) %>%
  count(COLLEGE_CLEAN, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

COURSE_CLASS_CT <- DFTREND %>%
  count(COURSE_CLASS_CLEAN, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

COURSE_RECODE_CT <- DFTREND %>%
  count(COURSE_RECODE_CLEAN, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

PERM_REGION_CT <- DFTREND %>%
  filter(!is.na(PERM_REGION)) %>%
  count(PERM_REGION, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

LOCAL_PROV_CT <- DFTREND %>%
  filter(!is.na(LOCAL_PROVINCE)) %>%
  count(LOCAL_PROVINCE, sort = TRUE, name = "n") %>%
  mutate(pct = round(n / sum(n) * 100, 2))

TOP15_NAT <- NAT_CT %>% slice_head(n = 15) %>% pull(NATIONALITY_CLEAN)
TOP10_CENTER <- CENTER_CT %>% slice_head(n = 10) %>% pull(CENTER_CLEAN)
TOP15_COLLEGE <- COLLEGE_CT %>% slice_head(n = 15) %>% pull(COLLEGE_CLEAN)
TOP12_COURSE <- COURSE_RECODE_CT %>% slice_head(n = 12) %>% pull(COURSE_RECODE_CLEAN)

NAT_SUMMARY <- DFTREND %>%
  filter(!is.na(NATIONALITY_CLEAN)) %>%
  group_by(NATIONALITY_CLEAN) %>%
  summarise(
    n = n(),
    median_percentile = safe_median(NMSPERnum),
    mean_percentile = safe_mean(NMSPERnum),
    median_gps = safe_median(NMSGPS),
    top_bin_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    bottom_bin_pct = round(mean(PercentileBin %in% c("B1", "B2", "B3"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(n), desc(median_percentile))

CENTER_SUMMARY <- DFTREND %>%
  filter(!is.na(CENTER_CLEAN)) %>%
  group_by(CENTER_CLEAN) %>%
  summarise(
    n = n(),
    median_percentile = safe_median(NMSPERnum),
    mean_percentile = safe_mean(NMSPERnum),
    median_gps = safe_median(NMSGPS),
    top_bin_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(n), desc(median_percentile))

SCHOOL_SUMMARY <- DFTREND %>%
  group_by(SCHOOL_TYPE) %>%
  summarise(
    n = n(),
    median_percentile = safe_median(NMSPERnum),
    mean_percentile = safe_mean(NMSPERnum),
    median_gps = safe_median(NMSGPS),
    top_bin_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    bottom_bin_pct = round(mean(PercentileBin %in% c("B1", "B2", "B3"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(n), desc(median_percentile))

COLLEGE_SUMMARY <- DFTREND %>%
  filter(!is.na(COLLEGE_CLEAN)) %>%
  group_by(COLLEGE_CLEAN) %>%
  summarise(
    n = n(),
    median_percentile = safe_median(NMSPERnum),
    mean_percentile = safe_mean(NMSPERnum),
    median_gps = safe_median(NMSGPS),
    top_bin_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(n), desc(median_percentile))

COURSE_RECODE_SUMMARY <- DFTREND %>%
  group_by(COURSE_RECODE_CLEAN) %>%
  summarise(
    n = n(),
    median_percentile = safe_median(NMSPERnum),
    mean_percentile = safe_mean(NMSPERnum),
    median_gps = safe_median(NMSGPS),
    top_bin_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    bottom_bin_pct = round(mean(PercentileBin %in% c("B1", "B2", "B3"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(n), desc(median_percentile))

COURSE_CLASS_SUMMARY <- DFTREND %>%
  group_by(COURSE_CLASS_CLEAN) %>%
  summarise(
    n = n(),
    median_percentile = safe_median(NMSPERnum),
    mean_percentile = safe_mean(NMSPERnum),
    median_gps = safe_median(NMSGPS),
    top_bin_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(n), desc(median_percentile))
```

<div class="callout-info">
<b>Overview.</b> This report is a foreigner-only NMAT dashboard built from the real foreigner CSV. It preserves the same dashboard style as the main report, but all computations here are restricted to fields that actually exist in the foreigner dataset.
</div>

## Executive Summary

```{r exsum_metrics, results='asis'}
n_total <- nrow(DFTREND)
n_years <- n_distinct(DFTREND$Year)
n_nat <- n_distinct(na.omit(DFTREND$NATIONALITY_CLEAN))
n_colleges <- n_distinct(na.omit(DFTREND$COLLEGE_CLEAN))
med_pct <- safe_median(DFTREND$NMSPERnum)
mean_pct <- safe_mean(DFTREND$NMSPERnum)
med_gps <- safe_median(DFTREND$NMSGPS)
top_nat <- NAT_CT %>% slice(1) %>% pull(NATIONALITY_CLEAN)
top_nat_pct <- NAT_CT %>% slice(1) %>% pull(pct)
top_center <- CENTER_CT %>% slice(1) %>% pull(CENTER_CLEAN)
top_center_pct <- CENTER_CT %>% slice(1) %>% pull(pct)

browsable(
  metric_row(
    metric_card(format(n_total, big.mark = ","), "Foreign examinee rows", "#7b1fa2"),
    metric_card(n_years, "Years covered (2006-2018)", "#3949ab"),
    metric_card(n_nat, "Clean nationality labels", "#00897b"),
    metric_card(n_colleges, "Distinct colleges", "#6d4c41"),
    metric_card(med_pct, "Median percentile", "#c62828"),
    metric_card(mean_pct, "Mean percentile", "#ef6c00"),
    metric_card(med_gps, "Median GPS", "#2e7d32"),
    metric_card(paste0(top_center, " (", top_center_pct, "%)"), "Largest testing center", "#1565c0")
  )
)
```

```{r exsum_callout, results='asis'}
pct_perm_overseas <- round(mean(DFTREND$IS_OVERSEAS_PERM, na.rm = TRUE) * 100, 2)
pct_local_overseas <- round(mean(DFTREND$IS_OVERSEAS_LOCAL, na.rm = TRUE) * 100, 2)
cat(sprintf(
  '<div class="callout-info"><b>What this summary emphasizes.</b> This dashboard reports foreign examinee volume, score distribution, nationality mix, testing-center concentration, school background, and subtest profiles. It also distinguishes <b>nationality</b> from address/location fields: %.2f%% have permanent region coded as Overseas, while %.2f%% have local province coded as Overseas.</div>',
  pct_perm_overseas, pct_local_overseas
))
```

```{r exsum_trend, fig.height=8}
p1 <- plot_ly(YEAR_SUM, x = ~Year) %>%
  add_ribbons(ymin = ~perq25, ymax = ~perq75, name = "IQR percentile",
              fillcolor = "rgba(214,39,40,0.18)", line = list(color = "rgba(214,39,40,0.05)"),
              hoverinfo = "skip", showlegend = FALSE) %>%
  add_lines(y = ~permedian, name = "Median percentile", line = list(color = "#d62728", width = 3)) %>%
  layout(title = "Percentile by year", yaxis = list(title = "Percentile"), xaxis = list(title = "Year"))

p2 <- plot_ly(YEAR_SUM, x = ~Year) %>%
  add_ribbons(ymin = ~gpsq25, ymax = ~gpsq75, name = "IQR GPS",
              fillcolor = "rgba(31,119,180,0.18)", line = list(color = "rgba(31,119,180,0.05)"),
              hoverinfo = "skip", showlegend = FALSE) %>%
  add_lines(y = ~gpsmedian, name = "Median GPS", line = list(color = "#1f77b4", width = 3)) %>%
  layout(title = "GPS by year", yaxis = list(title = "GPS"), xaxis = list(title = "Year"))

p3 <- plot_ly(YEAR_SUM, x = ~Year) %>%
  add_lines(y = ~aptmedian, name = "Aptitude", line = list(color = "#2ca02c", width = 3)) %>%
  add_lines(y = ~samedian, name = "SA", line = list(color = "#ff7f0e", width = 3)) %>%
  layout(title = "Aptitude and SA by year", yaxis = list(title = "Score"), xaxis = list(title = "Year"))

p4 <- plot_ly(YEAR_SUM, x = ~Year, y = ~n, type = "bar", name = "Examinees",
              marker = list(color = "#7b1fa2")) %>%
  layout(title = "Examinee count by year", yaxis = list(title = "Count"), xaxis = list(title = "Year"))

subplot(p1, p2, p3, p4, nrows = 2, titleX = TRUE, titleY = TRUE) %>%
  layout(height = 720, hovermode = "x unified")
```

```{r exsum_year_table}
dtexport(YEAR_SUM, "Annual Foreign Examinee Summary by Year")
```

```{r exsum_comp_pies, fig.height=5}
p_nat <- NAT_CT %>%
  mutate(group = ifelse(row_number() <= 10, NATIONALITY_CLEAN, "Others")) %>%
  group_by(group) %>%
  summarise(n = sum(n), .groups = "drop")

p_school <- SCHOOL_CT
p_course <- COURSE_RECODE_CT %>% mutate(group = ifelse(row_number() <= 8, COURSE_RECODE_CLEAN, "Others")) %>%
  group_by(group) %>% summarise(n = sum(n), .groups = "drop")

fig1 <- plot_ly(p_nat, labels = ~group, values = ~n, type = "pie") %>% layout(title = "Nationality composition (Top 10 + Others)")
fig2 <- plot_ly(p_school, labels = ~SCHOOL_TYPE, values = ~n, type = "pie") %>% layout(title = "School type composition")
fig3 <- plot_ly(p_course, labels = ~group, values = ~n, type = "pie") %>% layout(title = "Course recode composition")

subplot(fig1, fig2, fig3, nrows = 1, margin = 0.04) %>% layout(height = 420, showlegend = TRUE)
```

```{r exsum_summary_table}
SUMMARY_INDICATORS <- tibble(
  Indicator = c(
    "Median percentile",
    "Mean percentile",
    "Median GPS",
    "Top-bin share (B8-B10)",
    "Bottom-bin share (B1-B3)",
    "Rows with percentile",
    "Rows with all standardized subtests",
    "Permanent region coded Overseas"
  ),
  Value = c(
    safe_median(DFTREND$NMSPERnum),
    safe_mean(DFTREND$NMSPERnum),
    safe_median(DFTREND$NMSGPS),
    round(mean(DFTREND$PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    round(mean(DFTREND$PercentileBin %in% c("B1", "B2", "B3"), na.rm = TRUE) * 100, 2),
    sum(!is.na(DFTREND$NMSPERnum)),
    sum(DFTREND$HasAllStdSubtests, na.rm = TRUE),
    paste0(round(mean(DFTREND$IS_OVERSEAS_PERM, na.rm = TRUE) * 100, 2), "%")
  )
)

dtexport(SUMMARY_INDICATORS, "Executive Summary Indicators")
```

<hr class="break">

## Data Integrity

<div class="callout-info">
<b>Integrity focus.</b> This section documents completeness, duplicate application checks, and the cleaning of fragmented raw labels such as nationality and testing center. Use the cleaned fields for official tables and figures.
</div>

```{r data_integrity_metrics, results='asis'}
dup_apps <- DFTREND %>% filter(!is.na(NMA_AppNo)) %>% count(NMA_AppNo) %>% filter(n > 1) %>% nrow()
rows_with_pct <- sum(!is.na(DFTREND$NMSPERnum))
rows_missing_pct <- sum(is.na(DFTREND$NMSPERnum))
nat_raw_n <- n_distinct(na.omit(DFTREND$NATIONALITY_RAW))
nat_clean_n <- n_distinct(na.omit(DFTREND$NATIONALITY_CLEAN))
center_raw_n <- n_distinct(na.omit(DFTREND$CENTER_RAW))
center_clean_n <- n_distinct(na.omit(DFTREND$CENTER_CLEAN))

browsable(
  metric_row(
    metric_card(nrow(DFTREND), "Rows in analysis subset", "#6a1b9a"),
    metric_card(ncol(DFTREND), "Columns available", "#283593"),
    metric_card(rows_with_pct, "Rows with percentile", "#00897b"),
    metric_card(rows_missing_pct, "Rows missing percentile", "#ef6c00"),
    metric_card(dup_apps, "Duplicate application numbers", "#c62828"),
    metric_card(paste0(nat_raw_n, " -> ", nat_clean_n), "Nationality labels: raw to clean", "#5d4037"),
    metric_card(paste0(center_raw_n, " -> ", center_clean_n), "Center labels: raw to clean", "#1565c0")
  )
)
```

```{r data_integrity_quality}
dtexport(QUALITY_TBL, "Data Quality Checks")
```

```{r data_integrity_missing}
dtexport(MISSING_TBL, "Missingness by Column")
```

```{r data_integrity_harmonization}
dtexport(HARMONIZATION_TBL, "Raw vs Clean Category Counts")
```

```{r data_integrity_nat_compare}
nat_compare <- list(
  raw = NAT_RAW_CT %>% slice_head(n = 25) %>% rename(label = NATIONALITY_RAW),
  clean = NAT_CT %>% slice_head(n = 25) %>% rename(label = NATIONALITY_CLEAN)
)

dtexport(nat_compare$raw, "Top 25 Raw Nationality Labels")
```

```{r data_integrity_nat_compare_clean}
dtexport(nat_compare$clean, "Top 25 Clean Nationality Labels")
```

```{r data_integrity_center_compare}
center_compare <- tibble(
  Raw_center_distinct = n_distinct(na.omit(DFTREND$CENTER_RAW)),
  Clean_center_distinct = n_distinct(na.omit(DFTREND$CENTER_CLEAN)),
  Example_variants = "Los Angeles  (US) / Los Angeles (Us) -> Los Angeles (US)"
)

dtexport(center_compare, "Testing Center Harmonization Notes")
```

```{r data_integrity_distributions}
dtexport(SCHOOL_CT, "School Type Distribution")
```

```{r data_integrity_course_recode}
dtexport(COURSE_RECODE_CT, "Course Recode Distribution")
```

```{r data_integrity_perm_region}
dtexport(PERM_REGION_CT, "Permanent Region Distribution")
```

<hr class="break">

## Performance Trends

<div class="callout-info">
<b>Interpretation.</b> These plots track score distributions over time using only measures available in the foreigner CSV: percentile, GPS, aptitude, SA, and standardized subtests.
</div>

```{r trends_boxplot_percentile, fig.height=5}
df_pct <- DFTREND %>% filter(!is.na(Year), !is.na(NMSPERnum))
plot_ly(
  df_pct,
  x = ~factor(Year),
  y = ~NMSPERnum,
  type = "box",
  boxpoints = FALSE,
  marker = list(color = "#d62728")
) %>%
  layout(
    title = "Percentile Rank by Year",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Percentile")
  )
```

```{r trends_boxplot_gps, fig.height=5}
df_gps <- DFTREND %>% filter(!is.na(Year), !is.na(NMSGPS))
plot_ly(
  df_gps,
  x = ~factor(Year),
  y = ~NMSGPS,
  type = "box",
  boxpoints = FALSE,
  marker = list(color = "#1f77b4")
) %>%
  layout(
    title = "GPS by Year",
    xaxis = list(title = "Year"),
    yaxis = list(title = "GPS")
  )
```

```{r trends_stats_tables}
PCT_BY_YEAR <- DFTREND %>%
  filter(!is.na(Year), !is.na(NMSPERnum)) %>%
  group_by(Year) %>%
  summarise(
    n = n(),
    median = safe_median(NMSPERnum),
    q25 = safe_q(NMSPERnum, 0.25),
    q75 = safe_q(NMSPERnum, 0.75),
    mean = safe_mean(NMSPERnum),
    .groups = "drop"
  )

GPS_BY_YEAR <- DFTREND %>%
  filter(!is.na(Year), !is.na(NMSGPS)) %>%
  group_by(Year) %>%
  summarise(
    n = n(),
    median = safe_median(NMSGPS),
    q25 = safe_q(NMSGPS, 0.25),
    q75 = safe_q(NMSGPS, 0.75),
    mean = safe_mean(NMSGPS),
    .groups = "drop"
  )

dtexport(PCT_BY_YEAR, "Percentile Statistics by Year")
```

```{r trends_gps_table}
dtexport(GPS_BY_YEAR, "GPS Statistics by Year")
```

```{r trends_kw}
kw_res <- kwtable(
  DFTREND,
  "Year",
  list(
    NMSPERnum = "Percentile Rank",
    NMSGPS = "GPS",
    NMSAPT = "Aptitude",
    NMSSA = "SA",
    SubtestCompositeStd = "Subtest Composite Mean"
  )
)

dtexport(kw_res, "Kruskal-Wallis Tests Across Years")
```

```{r trends_subtest_heatmap, fig.height=6}
year_subtests <- subtestmeans(DFTREND, "Year")
subtestheatmap(year_subtests, "Mean Standardized Subtests by Year")
```

```{r trends_subtest_table}
dtexport(year_subtests, "Mean Standardized Subtests by Year")
```

<hr class="break">

## Score Bins by Year

<div class="callout-info">
<b>Bin framework.</b> Percentiles are grouped into B1-B10 using the same decile-style cut structure used in the main dashboard: B1 = 0-9, ..., B10 = 90-100.
</div>

```{r bins_year_heatmap, fig.height=6}
res_year_bin <- pctcross(DFTREND, "Year", "PercentileBin", BINORDER)
mat_year_bin <- as.matrix(res_year_bin$pct[, BINORDER, drop = FALSE])
rownames(mat_year_bin) <- as.character(res_year_bin$pct$Year)

plot_ly(
  z = mat_year_bin,
  x = BINORDER,
  y = rownames(mat_year_bin),
  type = "heatmap",
  colorscale = "YlOrRd",
  hovertemplate = "Year: %{y}<br>Bin: %{x}<br>Row %: %{z}<extra></extra>"
) %>%
  layout(
    title = "Percentile Bin by Year (Row Percentages)",
    xaxis = list(title = "Bin"),
    yaxis = list(title = "Year")
  )
```

```{r bins_year_pct_table}
dtexport(res_year_bin$pct %>% mutate(Year = as.character(Year)), "Percentile Bin by Year - Row Percentages")
```

```{r bins_year_count_table}
dtexport(res_year_bin$counts %>% mutate(Year = as.character(Year)), "Percentile Bin by Year - Counts")
```

```{r bins_year_stacked, fig.height=5}
long_year_bin <- res_year_bin$pct %>%
  mutate(Year = as.integer(Year)) %>%
  pivot_longer(-Year, names_to = "Bin", values_to = "Pct") %>%
  mutate(Bin = factor(Bin, levels = BINORDER, ordered = TRUE))

plot_ly(
  long_year_bin,
  x = ~Year,
  y = ~Pct,
  color = ~Bin,
  colors = PALBIN,
  type = "bar"
) %>%
  layout(
    barmode = "stack",
    title = "Bin Composition by Year",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Percent")
  )
```

```{r bins_year_top_bottom, fig.height=5}
tb_year <- res_year_bin$pct %>%
  mutate(
    Year = as.integer(Year),
    Top_B8_B10 = rowSums(dplyr::select(., any_of(c("B8", "B9", "B10"))), na.rm = TRUE),
    Bottom_B1_B3 = rowSums(dplyr::select(., any_of(c("B1", "B2", "B3"))), na.rm = TRUE)
  ) %>%
  select(Year, Top_B8_B10, Bottom_B1_B3)

plot_ly(tb_year, x = ~Year) %>%
  add_lines(y = ~Top_B8_B10, name = "Top bins B8-B10", line = list(color = "#2e7d32", width = 3)) %>%
  add_lines(y = ~Bottom_B1_B3, name = "Bottom bins B1-B3", line = list(color = "#c62828", width = 3)) %>%
  layout(
    title = "Top vs Bottom Bin Share by Year",
    xaxis = list(title = "Year"),
    yaxis = list(title = "Percent")
  )
```

```{r bins_year_top_bottom_table}
dtexport(tb_year, "Top vs Bottom Bin Share by Year")
```

<hr class="break">

## Nationality Analysis

<div class="callout-info">
<b>Important.</b> All reporting charts in this section use <b>NATIONALITY_CLEAN</b>, not the raw nationality field, so that obvious duplicates and spelling variants do not fragment the counts.
</div>

```{r nat_metrics, results='asis'}
top5_share <- NAT_CT %>% slice_head(n = 5) %>% summarise(v = round(sum(pct), 2)) %>% pull(v)
top1_nat <- NAT_CT %>% slice(1) %>% pull(NATIONALITY_CLEAN)
top1_nat_pct <- NAT_CT %>% slice(1) %>% pull(pct)

browsable(
  metric_row(
    metric_card(n_distinct(na.omit(DFTREND$NATIONALITY_RAW)), "Raw nationality labels", "#5e35b1"),
    metric_card(n_distinct(na.omit(DFTREND$NATIONALITY_CLEAN)), "Clean nationality labels", "#3949ab"),
    metric_card(paste0(top1_nat, " (", top1_nat_pct, "%)"), "Largest nationality group", "#00897b"),
    metric_card(paste0(top5_share, "%"), "Top 5 nationality share", "#ef6c00")
  )
)
```

```{r nat_top15_bar, fig.height=6}
nat_bar <- NAT_CT %>% slice_head(n = 15)

plot_ly(
  nat_bar,
  x = ~n,
  y = ~reorder(NATIONALITY_CLEAN, n),
  type = "bar",
  orientation = "h",
  marker = list(color = "#7b1fa2"),
  text = ~paste0(pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title = "Top 15 Clean Nationalities by Count",
    xaxis = list(title = "Examinees"),
    yaxis = list(title = "")
  )
```

```{r nat_top15_table}
dtexport(nat_bar, "Top 15 Clean Nationalities by Count")
```

```{r nat_bin_heatmap, fig.height=6}
df_nat_top <- DFTREND %>% filter(NATIONALITY_CLEAN %in% TOP15_NAT)
res_nat_bin <- pctcross(df_nat_top, "NATIONALITY_CLEAN", "PercentileBin", BINORDER)
mat_nat_bin <- as.matrix(res_nat_bin$pct[, BINORDER, drop = FALSE])
rownames(mat_nat_bin) <- as.character(res_nat_bin$pct$NATIONALITY_CLEAN)

plot_ly(
  z = mat_nat_bin,
  x = BINORDER,
  y = rownames(mat_nat_bin),
  type = "heatmap",
  colorscale = "YlOrRd",
  hovertemplate = "Nationality: %{y}<br>Bin: %{x}<br>Row %: %{z}<extra></extra>"
) %>%
  layout(
    title = "Top 15 Nationalities by Percentile Bin (Row Percentages)",
    xaxis = list(title = "Bin"),
    yaxis = list(title = "")
  )
```

```{r nat_bin_table}
dtexport(res_nat_bin$pct, "Top 15 Nationalities by Percentile Bin - Row Percentages")
```

```{r nat_topbin_bar, fig.height=6}
nat_topbin <- NAT_SUMMARY %>%
  filter(n >= 20) %>%
  slice_head(n = 20) %>%
  arrange(top_bin_pct)

plot_ly(
  nat_topbin,
  x = ~top_bin_pct,
  y = ~reorder(NATIONALITY_CLEAN, top_bin_pct),
  type = "bar",
  orientation = "h",
  marker = list(color = "#2e7d32"),
  text = ~paste0(top_bin_pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title = "Top-Bin Share (B8-B10) by Nationality, n >= 20",
    xaxis = list(title = "Percent in B8-B10"),
    yaxis = list(title = "")
  )
```

```{r nat_boxplot_percentile, fig.height=6}
nat_box <- DFTREND %>%
  filter(NATIONALITY_CLEAN %in% TOP15_NAT, !is.na(NMSPERnum))

plot_ly(
  nat_box,
  x = ~NATIONALITY_CLEAN,
  y = ~NMSPERnum,
  type = "box",
  boxpoints = FALSE,
  marker = list(color = "#8e24aa")
) %>%
  layout(
    title = "Percentile Rank by Top 15 Nationalities",
    xaxis = list(title = ""),
    yaxis = list(title = "Percentile")
  )
```

```{r nat_summary_table}
dtexport(NAT_SUMMARY, "Nationality Summary Table")
```

<hr class="break">

## Testing Center Analysis

<div class="callout-info">
<b>Reading tip.</b> Testing center is not the same as nationality. This section shows where foreign examinees actually sat the exam.
</div>

```{r center_top_bar, fig.height=6}
center_bar <- CENTER_CT %>% slice_head(n = 15)

plot_ly(
  center_bar,
  x = ~n,
  y = ~reorder(CENTER_CLEAN, n),
  type = "bar",
  orientation = "h",
  marker = list(color = "#1565c0"),
  text = ~paste0(pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title = "Top 15 Testing Centers by Count",
    xaxis = list(title = "Examinees"),
    yaxis = list(title = "")
  )
```

```{r center_top_table}
dtexport(center_bar, "Top 15 Testing Centers by Count")
```

```{r center_bin_heatmap, fig.height=6}
df_center_top <- DFTREND %>% filter(CENTER_CLEAN %in% TOP10_CENTER)
res_center_bin <- pctcross(df_center_top, "CENTER_CLEAN", "PercentileBin", BINORDER)
mat_center_bin <- as.matrix(res_center_bin$pct[, BINORDER, drop = FALSE])
rownames(mat_center_bin) <- as.character(res_center_bin$pct$CENTER_CLEAN)

plot_ly(
  z = mat_center_bin,
  x = BINORDER,
  y = rownames(mat_center_bin),
  type = "heatmap",
  colorscale = "Blues",
  hovertemplate = "Center: %{y}<br>Bin: %{x}<br>Row %: %{z}<extra></extra>"
) %>%
  layout(
    title = "Top 10 Testing Centers by Percentile Bin",
    xaxis = list(title = "Bin"),
    yaxis = list(title = "")
  )
```

```{r center_boxplot, fig.height=6}
center_box <- DFTREND %>% filter(CENTER_CLEAN %in% TOP10_CENTER, !is.na(NMSPERnum))

plot_ly(
  center_box,
  x = ~CENTER_CLEAN,
  y = ~NMSPERnum,
  type = "box",
  boxpoints = FALSE,
  marker = list(color = "#1e88e5")
) %>%
  layout(
    title = "Percentile Rank by Top 10 Testing Centers",
    xaxis = list(title = ""),
    yaxis = list(title = "Percentile")
  )
```

```{r center_summary_table}
dtexport(CENTER_SUMMARY, "Testing Center Summary Table")
```

```{r center_location_tables}
dtexport(PERM_REGION_CT, "Permanent Region Counts")
```

```{r center_local_province_table}
dtexport(LOCAL_PROV_CT, "Local Province Counts")
```

<hr class="break">

## School and College Analysis

<div class="callout-info">
<b>Scope.</b> This section replaces the Ultima university-type pages with analyses based on the fields that truly exist here: school type and college.
</div>

```{r school_type_pie, fig.height=5}
plot_ly(
  SCHOOL_CT,
  labels = ~SCHOOL_TYPE,
  values = ~n,
  type = "pie"
) %>%
  layout(title = "School Type Composition")
```

```{r school_type_heatmap, fig.height=5}
res_school_bin <- pctcross(DFTREND, "SCHOOL_TYPE", "PercentileBin", BINORDER)
mat_school_bin <- as.matrix(res_school_bin$pct[, BINORDER, drop = FALSE])
rownames(mat_school_bin) <- as.character(res_school_bin$pct$SCHOOL_TYPE)

plot_ly(
  z = mat_school_bin,
  x = BINORDER,
  y = rownames(mat_school_bin),
  type = "heatmap",
  colorscale = "YlGnBu",
  hovertemplate = "School type: %{y}<br>Bin: %{x}<br>Row %: %{z}<extra></extra>"
) %>%
  layout(
    title = "School Type by Percentile Bin",
    xaxis = list(title = "Bin"),
    yaxis = list(title = "")
  )
```

```{r school_type_topbin, fig.height=4}
plot_ly(
  SCHOOL_SUMMARY,
  x = ~top_bin_pct,
  y = ~reorder(SCHOOL_TYPE, top_bin_pct),
  type = "bar",
  orientation = "h",
  marker = list(color = "#2e7d32"),
  text = ~paste0(top_bin_pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title = "Top-Bin Share (B8-B10) by School Type",
    xaxis = list(title = "Percent in B8-B10"),
    yaxis = list(title = "")
  )
```

```{r school_type_boxplot, fig.height=5}
plot_ly(
  DFTREND %>% filter(!is.na(NMSPERnum)),
  x = ~SCHOOL_TYPE,
  y = ~NMSPERnum,
  type = "box",
  boxpoints = FALSE,
  marker = list(color = "#43a047")
) %>%
  layout(
    title = "Percentile Rank by School Type",
    xaxis = list(title = ""),
    yaxis = list(title = "Percentile")
  )
```

```{r school_summary_table}
dtexport(SCHOOL_SUMMARY, "School Type Summary Table")
```

```{r college_top_bar, fig.height=6}
college_bar <- COLLEGE_CT %>% slice_head(n = 20)

plot_ly(
  college_bar,
  x = ~n,
  y = ~reorder(COLLEGE_CLEAN, n),
  type = "bar",
  orientation = "h",
  marker = list(color = "#6d4c41"),
  text = ~paste0(pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title = "Top 20 Colleges by Count",
    xaxis = list(title = "Examinees"),
    yaxis = list(title = "")
  )
```

```{r college_top_table}
dtexport(college_bar, "Top 20 Colleges by Count")
```

```{r college_summary_filtered}
college_summary_top <- COLLEGE_SUMMARY %>% slice_head(n = 50)
dtexport(college_summary_top, "Top 50 Colleges - Score Summary")
```

<hr class="break">

## Course Background Analysis

<div class="callout-info">
<b>Fields used.</b> This section uses both <b>Course Classification</b> and <b>Course_recode</b>. The recoded course field is the main grouping field for performance charts.
</div>

```{r course_recode_pie, fig.height=5}
course_pie_tbl <- COURSE_RECODE_CT %>%
  mutate(group = ifelse(row_number() <= 10, COURSE_RECODE_CLEAN, "Others")) %>%
  group_by(group) %>%
  summarise(n = sum(n), .groups = "drop")

plot_ly(
  course_pie_tbl,
  labels = ~group,
  values = ~n,
  type = "pie"
) %>%
  layout(title = "Course Recode Composition (Top 10 + Others)")
```

```{r course_class_table}
dtexport(COURSE_CLASS_CT, "Course Classification Counts")
```

```{r course_recode_heatmap, fig.height=6}
df_course_top <- DFTREND %>% filter(COURSE_RECODE_CLEAN %in% TOP12_COURSE)
res_course_bin <- pctcross(df_course_top, "COURSE_RECODE_CLEAN", "PercentileBin", BINORDER)
mat_course_bin <- as.matrix(res_course_bin$pct[, BINORDER, drop = FALSE])
rownames(mat_course_bin) <- as.character(res_course_bin$pct$COURSE_RECODE_CLEAN)

plot_ly(
  z = mat_course_bin,
  x = BINORDER,
  y = rownames(mat_course_bin),
  type = "heatmap",
  colorscale = "YlOrRd",
  hovertemplate = "Course: %{y}<br>Bin: %{x}<br>Row %: %{z}<extra></extra>"
) %>%
  layout(
    title = "Top Course Recode Groups by Percentile Bin",
    xaxis = list(title = "Bin"),
    yaxis = list(title = "")
  )
```

```{r course_recode_boxplot, fig.height=6}
course_box <- DFTREND %>%
  filter(COURSE_RECODE_CLEAN %in% TOP12_COURSE, !is.na(NMSPERnum))

plot_ly(
  course_box,
  x = ~COURSE_RECODE_CLEAN,
  y = ~NMSPERnum,
  type = "box",
  boxpoints = FALSE,
  marker = list(color = "#fb8c00")
) %>%
  layout(
    title = "Percentile Rank by Top Course Recode Groups",
    xaxis = list(title = ""),
    yaxis = list(title = "Percentile")
  )
```

```{r course_recode_summary}
dtexport(COURSE_RECODE_SUMMARY, "Course Recode Summary Table")
```

```{r course_class_summary}
dtexport(COURSE_CLASS_SUMMARY, "Course Classification Summary Table")
```

<hr class="break">

## Subtest Profiles

<div class="callout-info">
<b>Reading tip.</b> These are mean standardized subtest scores, not percentiles. They help show relative strengths across verbal, reasoning, quantitative, perceptual, and science areas.
</div>

### By school type

```{r subtest_school_heatmap, fig.height=5}
sub_school <- subtestmeans(DFTREND, "SCHOOL_TYPE")
subtestheatmap(sub_school, "Mean Standardized Subtests by School Type")
```

```{r subtest_school_radar, fig.height=6}
radarplot_tbl(sub_school, "Radar Profile: School Type")
```

```{r subtest_school_table}
dtexport(sub_school, "Subtest Means by School Type")
```

### By course recode

```{r subtest_course_heatmap, fig.height=6}
sub_course <- subtestmeans(DFTREND %>% filter(COURSE_RECODE_CLEAN %in% TOP12_COURSE), "COURSE_RECODE_CLEAN")
subtestheatmap(sub_course, "Mean Standardized Subtests by Top Course Recode Groups")
```

```{r subtest_course_radar, fig.height=7}
radarplot_tbl(sub_course %>% slice_head(n = 8), "Radar Profile: Top Course Recode Groups")
```

```{r subtest_course_table}
dtexport(sub_course, "Subtest Means by Top Course Recode Groups")
```

### By nationality

```{r subtest_nat_heatmap, fig.height=6}
sub_nat <- subtestmeans(DFTREND %>% filter(NATIONALITY_CLEAN %in% TOP15_NAT), "NATIONALITY_CLEAN")
subtestheatmap(sub_nat, "Mean Standardized Subtests by Top 15 Nationalities")
```

```{r subtest_nat_radar, fig.height=7}
sub_nat_radar <- sub_nat %>% slice_head(n = 8)
radarplot_tbl(sub_nat_radar, "Radar Profile: Top Nationalities")
```

```{r subtest_nat_table}
dtexport(sub_nat, "Subtest Means by Top 15 Nationalities")
```

<hr class="break">

## Record Listings

<div class="callout-warning">
<b>Use with care.</b> These are record-level listings from the foreigner dataset. Keep them for internal analytic review, not for public release, unless your governance and privacy rules allow row-level disclosure.
</div>
