---
title: "NMAT Performance Dashboard, 2006–2018"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    theme: flatly
    code_folding: hide
    self_contained: true
    fig_width: 11
    fig_height: 7
---

```{r css, echo=FALSE, results="asis"}
cat('
<style>
.metric-row { display:flex; flex-wrap:wrap; gap:12px; margin:16px 0; }
.metric-card { background:#fff; border-left:4px solid #3c8dbc; border-radius:4px; padding:12px 18px; min-width:160px; box-shadow:0 1px 3px rgba(0,0,0,.12); }
.metric-value { font-size:1.6em; font-weight:700; color:#2c3e50; }
.metric-label { font-size:.82em; color:#7f8c8d; margin-top:2px; }
.callout-info { background:#e8f4f8; border-left:4px solid #3498db; padding:10px 16px; margin:12px 0; border-radius:2px; font-size:.92em; }
.callout-warning { background:#fff8e1; border-left:4px solid #f39c12; padding:10px 16px; margin:12px 0; border-radius:2px; font-size:.92em; }
.section-hdr { background:#f0f4ff; border-left:3px solid #1a237e; padding:8px 14px; margin:18px 0 10px 0; border-radius:2px; font-size:.95em; font-weight:600; color:#1a237e; }
hr.break { border:none; border-top:2px solid #e0e0e0; margin:22px 0; }
</style>
')
```

```{r setup, include=FALSE, cache=FALSE}
suppressPackageStartupMessages({
  library(arrow)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(plotly)
  library(DT)
  library(scales)
  library(RColorBrewer)
  library(htmltools)
  library(knitr)
})

HAS_DUNN <- requireNamespace("dunn.test", quietly = TRUE)

# ── CONSTANTS ──────────────────────────────────────────────
BIN_ORDER <- paste0("B", 1:10)
PLE_ORDER    <- c("Confirmed PLE passer", "No confirmed PLE match")

PAL_UNI <- c(
  "Public"       = "#1f77b4",
  "Private"      = "#ff7f0e",
  "Foreign"      = "#9467bd",
  "Not Specified"= "#7f7f7f"
)
PAL_COURSE <- c(
  "Medical & Allied"             = "#d62728",
  "Natural Sciences"             = "#2ca02c",
  "Social & Behavioral Sciences" = "#ff9800",
  "Education"                    = "#17becf",
  "Engineering & Technology"     = "#8c564b",
  "Other"                        = "#7f7f7f"
)
PAL_PLE <- c(
  "Confirmed PLE passer"    = "#2e7d32",
  "No confirmed PLE match"  = "#c62828"
)
PAL_BIN <- c(
  B1 = "#8B0000", B2 = "#B22222", B3 = "#D9534F",
  B4 = "#F0AD4E", B5 = "#FFD166", B6 = "#A0D468",
  B7 = "#66C2A5", B8 = "#41B6C4", B9 = "#2C7FB8", B10 = "#253494"
)

NUMERIC_COLS <- c(
  "Year","NMS_PER_num","NMS_GPS","NMS_APT","NMS_SA",
  "NMS_VCss","NMS_IRss","NMS_Qss","NMS_PAss",
  "NMS_BIOss","NMS_PHYss","NMS_SSCss","NMS_CHEMss",
  "TotalRawScoreTRUE","PartIRawScoreTRUE","PartIIRawScoreTRUE",
  "Raw_Verbal","Raw_InductiveReasoning","Raw_Quantitative","Raw_PerceptualAcuity",
  "Raw_Biology","Raw_Physics","Raw_SocialScience","Raw_Chemistry",
  "PLE_YEAR_PASSED","PLE_YEAR_GAP","PLE_MATCH_CONFIDENCE"
)

SUBTEST_STD <- c(
  Verbal       = "NMS_VCss",
  Inductive    = "NMS_IRss",
  Quantitative = "NMS_Qss",
  Perceptual   = "NMS_PAss",
  Biology      = "NMS_BIOss",
  Physics      = "NMS_PHYss",
  Social       = "NMS_SSCss",
  Chemistry    = "NMS_CHEMss"
)
SUBTEST_RAW <- c(
  Verbal       = "Raw_Verbal",
  Inductive    = "Raw_InductiveReasoning",
  Quantitative = "Raw_Quantitative",
  Perceptual   = "Raw_PerceptualAcuity",
  Biology      = "Raw_Biology",
  Physics      = "Raw_Physics",
  Social       = "Raw_SocialScience",
  Chemistry    = "Raw_Chemistry"
)

# ── DATA LOADING ───────────────────────────────────────────
find_parquet <- function() {
  candidates <- c(
    "../../dataset/NMAT_Ultima.parquet",
    "../dataset/NMAT_Ultima.parquet",
    "dataset/NMAT_Ultima.parquet",
    "NMAT_Ultima.parquet"
  )
  for (p in candidates) if (file.exists(p)) return(p)
  stop("Cannot find NMAT_Ultima.parquet — check working directory.")
}

load_and_prep <- function() {
  df <- arrow::read_parquet(find_parquet())

  for (col in intersect(NUMERIC_COLS, names(df)))
    df[[col]] <- suppressWarnings(as.numeric(df[[col]]))

  df$Year <- as.integer(df$Year)

  for (flag in c("IS_BEST_NMAT_RECORD","IS_PLE_ANALYSIS_SAFE",
                 "IS_PLE_PASSER","HasTRUErawScores","AllRawComponentsPresent")) {
    if (flag %in% names(df))
      df[[flag]] <- as.logical(df[[flag]])
  }

  if ("UNI_TYPE"    %in% names(df)) df$UNI_TYPE    <- ifelse(is.na(df$UNI_TYPE), "Not Specified", df$UNI_TYPE)
  if ("UNI_LOCATION"%in% names(df)) df$UNI_LOCATION<- ifelse(is.na(df$UNI_LOCATION), "Unknown", df$UNI_LOCATION)
  if ("CourseGroup" %in% names(df)) df$CourseGroup  <- ifelse(is.na(df$CourseGroup),  "Unknown",      df$CourseGroup)

  BIN_ORDER <- paste0("B", 1:10)
  # Back-compat: rename old column if present
  if ("PercentileDecile" %in% colnames(df) && !("PercentileBin" %in% colnames(df))) {
    df <- df %>% rename(PercentileBin = PercentileDecile)
  }
  if (!("PercentileBin" %in% colnames(df))) {
    df$PercentileBin <- cut(
      suppressWarnings(as.numeric(df$NMS_PER_num)),
      breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 101),
      labels = BIN_ORDER,
      right = FALSE,
      include.lowest = TRUE
    )
  }
  df$PercentileBin <- factor(df$PercentileBin, levels = BIN_ORDER, ordered = TRUE)

  if (!"SEX_CLEAN" %in% names(df)) {
    src <- if ("SEX" %in% names(df)) "SEX" else if ("NMASex" %in% names(df)) "NMASex" else NA
    if (!is.na(src)) {
      s <- as.character(df[[src]])
      s <- ifelse(s == "1","Male", ifelse(s == "2","Female", s))
      df$SEX_CLEAN <- ifelse(s %in% c("Male","Female"), s, NA_character_)
    } else df$SEX_CLEAN <- NA_character_
  }

  if (!"IS_BOARD_OBSERVABLE_COHORT" %in% names(df))
    df$IS_BOARD_OBSERVABLE_COHORT <- !is.na(df$Year) & df$Year <= 2014
  else
    df$IS_BOARD_OBSERVABLE_COHORT <- as.logical(df$IS_BOARD_OBSERVABLE_COHORT)

  if (!"PLE_STATUS_LABEL" %in% names(df)) {
    df$PLE_STATUS_LABEL <- ifelse(
      !is.na(df$IS_PLE_ANALYSIS_SAFE) & df$IS_PLE_ANALYSIS_SAFE,
      "Confirmed PLE passer", "No confirmed PLE match")
  }
  df$PLE_STATUS_LABEL <- factor(df$PLE_STATUS_LABEL, levels = PLE_ORDER)

  df$HAS_CONFIRMED_PLE <- !is.na(df$IS_PLE_ANALYSIS_SAFE) & df$IS_PLE_ANALYSIS_SAFE
  df
}

DF_RAW <- load_and_prep()
DF_BEST       <- DF_RAW  %>% filter(!is.na(IS_BEST_NMAT_RECORD), IS_BEST_NMAT_RECORD)
DF_TREND      <- DF_RAW  %>% filter(!is.na(Year), Year >= 2006, Year <= 2018)
DF_BESTTREND  <- DF_BEST %>% filter(!is.na(Year), Year >= 2006, Year <= 2018)
DF_BESTOBS    <- DF_BESTTREND %>% filter(IS_BOARD_OBSERVABLE_COHORT)
DF_UNI        <- DF_BESTTREND %>% filter(UNI_TYPE %in% c("Public","Private","Foreign"))
DF_UNIOBS     <- DF_BESTOBS   %>% filter(UNI_TYPE %in% c("Public","Private","Foreign"))

# ── HELPER FUNCTIONS ───────────────────────────────────────

`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

apply_filters <- function(df, yrs, courses, sexes) {
  if (!is.null(yrs)     && length(yrs)     > 0) df <- df %>% filter(Year       %in% yrs)
  if (!is.null(courses) && length(courses) > 0) df <- df %>% filter(CourseGroup %in% courses)
  if (!is.null(sexes)   && length(sexes)   > 0) df <- df %>% filter(SEX_CLEAN   %in% sexes)
  df
}

sanitize_for_dt <- function(df) {
  df <- as.data.frame(df, stringsAsFactors = FALSE)
  for (col in names(df)) {
    x <- df[[col]]
    if (is.list(x)) {
      df[[col]] <- vapply(x, function(v) {
        if (is.null(v) || length(v) == 0) NA_character_
        else paste(as.character(v), collapse = "; ")
      }, character(1))
    } else if (inherits(x, c("arrow", "ArrowArray", "ChunkedArray"))) {
      df[[col]] <- as.vector(x)
    }
    if (is.factor(df[[col]])) df[[col]] <- as.character(df[[col]])
  }
  df
}

dt_export <- function(data, caption = NULL) {
  data <- sanitize_for_dt(data)
  datatable(data,
    caption = if (!is.null(caption))
      tags$caption(style = "font-weight:bold;font-size:13px;text-align:left", caption),
    extensions = c("Buttons","Scroller"),
    options    = list(
      dom     = "Blfrtip",
      buttons = list(
        list(extend = "copy",  text = "Copy"),
        list(extend = "csv",   text = "CSV",
             filename = gsub("[^A-Za-z0-9_]","_", caption %||% "NMAT_export")),
        list(extend = "excel", text = "Excel",
             filename = gsub("[^A-Za-z0-9_]","_", caption %||% "NMAT_export"),
             title    = caption %||% "NMAT Export"),
        list(extend = "print", text = "Print")
      ),
      scrollX    = TRUE,
      scrollY    = "320px",
      scroller   = TRUE,
      pageLength = 25,
      lengthMenu = list(c(25, 50, 100, -1), c("25", "50", "100", "All"))
    ),
    rownames = FALSE,
    class    = "display compact stripe"
  )
}

yearly_summary <- function(df) {
  df %>%
    filter(!is.na(Year)) %>%
    group_by(Year) %>%
    summarise(
      n            = n(),
      raw_median   = median(TotalRawScoreTRUE, na.rm = TRUE),
      raw_q25      = quantile(TotalRawScoreTRUE, .25, na.rm = TRUE),
      raw_q75      = quantile(TotalRawScoreTRUE, .75, na.rm = TRUE),
      part1_median = median(PartIRawScoreTRUE,   na.rm = TRUE),
      part2_median = median(PartIIRawScoreTRUE,  na.rm = TRUE),
      per_median   = median(NMS_PER_num,         na.rm = TRUE),
      per_q25      = quantile(NMS_PER_num, .25,  na.rm = TRUE),
      per_q75      = quantile(NMS_PER_num, .75,  na.rm = TRUE),
      .groups      = "drop"
    )
}

kw_table <- function(df, group_col, col_map) {
  bind_rows(lapply(names(col_map), function(col) {
    if (!col %in% names(df)) return(NULL)
    sub <- df %>% filter(!is.na(.data[[col]]), !is.na(.data[[group_col]]))
    grps <- split(sub[[col]], sub[[group_col]])
    grps <- Filter(function(g) length(g) > 1, grps)
    if (length(grps) < 2) return(NULL)
    tryCatch({
      kt <- kruskal.test(grps)
      k  <- length(grps); n <- nrow(sub)
      eta2 <- max(0, (kt$statistic - k + 1) / (n - k))
      tibble(Score = col_map[[col]], H = round(kt$statistic, 3),
             p_value = ifelse(kt$p.value < .001, "<0.001", round(kt$p.value, 4)),
             eta_squared = round(eta2, 4))
    }, error = function(e) NULL)
  }))
}

mw_ple <- function(df, col_map) {
  g1 <- df %>% filter(PLE_STATUS_LABEL == "Confirmed PLE passer")
  g2 <- df %>% filter(PLE_STATUS_LABEL == "No confirmed PLE match")
  bind_rows(lapply(names(col_map), function(col) {
    if (!col %in% names(df)) return(NULL)
    a <- g1[[col]][!is.na(g1[[col]])]; b <- g2[[col]][!is.na(g2[[col]])]
    if (length(a) < 5 || length(b) < 5) return(NULL)
    tryCatch({
      wt <- wilcox.test(a, b, exact = FALSE)
      W  <- as.numeric(wt$statistic)
      r  <- 1 - 2 * W / (length(a) * length(b))
      tibble(Score = col_map[[col]], U_stat = round(W, 1),
             p_value = ifelse(wt$p.value < .001, "<0.001", round(wt$p.value, 4)),
             effect_r = round(r, 4),
             Confirmed_median = round(median(a), 2),
             NoMatch_median   = round(median(b), 2))
    }, error = function(e) NULL)
  }))
}

pct_cross <- function(df, row_col, col_col, col_order = NULL) {
  ct <- df %>%
    filter(!is.na(.data[[row_col]]), !is.na(.data[[col_col]])) %>%
    count(.data[[row_col]], .data[[col_col]]) %>%
    pivot_wider(names_from  = all_of(col_col),
                values_from = n, values_fill = 0L)
  if (!is.null(col_order)) {
    missing <- setdiff(col_order, names(ct))
    for (m in missing) ct[[m]] <- 0L
    keep <- c(row_col, intersect(col_order, names(ct)))
    ct   <- ct[, keep, drop = FALSE]
  }
  totals <- rowSums(ct[, -1, drop = FALSE])
  pct    <- ct
  pct[, -1] <- round(sweep(ct[, -1, drop = FALSE], 1, pmax(totals, 1), "/") * 100, 2)
  list(counts = ct, pct = pct)
}

make_sankey <- function(flow_df, src_col, tgt_col, value_col = "n",
                        title = "") {
  src_nodes <- unique(as.character(flow_df[[src_col]]))
  tgt_nodes <- unique(as.character(flow_df[[tgt_col]]))
  nodes     <- c(src_nodes, tgt_nodes)
  idx       <- setNames(seq_along(nodes) - 1L, nodes)

  srcs <- as.integer(idx[as.character(flow_df[[src_col]])])
  tgts <- as.integer(idx[as.character(flow_df[[tgt_col]])])
  vals <- as.numeric(flow_df[[value_col]])

  plot_ly(
    type        = "sankey",
    arrangement = "snap",
    node = list(label = nodes, pad = 16, thickness = 18,
                line = list(color = "black", width = 0.4)),
    link = list(source = srcs, target = tgts, value = vals)
  ) %>% layout(title = title, height = 520, font = list(size = 12))
}

subtest_means <- function(df, group_col, use_std = TRUE) {
  cols <- if (use_std) SUBTEST_STD else SUBTEST_RAW
  # cols is a named vector: names = display labels ("Verbal"), values = df column names ("NMS_VCss")
  mask      <- unname(cols) %in% names(df)
  col_vals  <- unname(cols)[mask]   # actual column names present in df
  col_nms   <- names(cols)[mask]    # display names to use in output
  if (length(col_vals) == 0 || !group_col %in% names(df)) return(NULL)
  result <- df %>%
    dplyr::filter(!is.na(.data[[group_col]])) %>%
    dplyr::group_by(.data[[group_col]]) %>%
    dplyr::summarise(dplyr::across(all_of(col_vals), ~ round(mean(.x, na.rm = TRUE), 2)),
                     .groups = "drop")
  # Rename output columns from actual names ("NMS_VCss") to display names ("Verbal").
  # Using base R to avoid tidyselect named-vector semantics that vary by version.
  for (i in seq_along(col_vals)) {
    j <- which(names(result) == col_vals[i])
    if (length(j) > 0L) names(result)[j] <- col_nms[i]
  }
  result
}

subtest_heatmap <- function(tbl, title) {
  if (is.null(tbl) || nrow(tbl) == 0) return(plot_ly() %>% layout(title="No data"))
  g_col   <- names(tbl)[1]
  mat_raw <- as.data.frame(tbl)[, -1, drop = FALSE]
  mat     <- matrix(
    as.numeric(as.matrix(mat_raw)),
    nrow     = nrow(tbl),
    ncol     = ncol(mat_raw),
    dimnames = list(as.character(tbl[[g_col]]), colnames(mat_raw))
  )
  plot_ly(z=mat, x=colnames(mat), y=rownames(mat), type="heatmap",
          colorscale="RdBu", reversescale=TRUE,
          texttemplate="%{z:.1f}", showscale=TRUE) %>%
    layout(title=title, xaxis=list(title="Subtest"), yaxis=list(title=""))
}

radar_plot <- function(tbl, title) {
  if (is.null(tbl) || nrow(tbl) == 0) return(plot_ly() %>% layout(title="No data"))
  g_col <- names(tbl)[1]
  cats  <- names(tbl)[-1]
  fig   <- plot_ly(type="scatterpolar", fill="toself")
  for (i in seq_len(nrow(tbl))) {
    vals <- as.numeric(unlist(tbl[i, -1], use.names = FALSE))
    fig  <- fig %>% add_trace(
      r = c(vals, vals[1]), theta = c(cats, cats[1]), name = tbl[[g_col]][i])
  }
  fig %>% layout(title=title,
                 polar=list(radialaxis=list(visible=TRUE, range=c(450,550))),
                 showlegend=TRUE)
}

make_flow <- function(df, src, tgt, src_order=NULL, tgt_order=NULL) {
  ct <- df %>% filter(!is.na(.data[[src]]), !is.na(.data[[tgt]])) %>%
    count(.data[[src]], .data[[tgt]])
  names(ct)[1:2] <- c("source","target")
  if (!is.null(src_order)) {
    ct$source <- factor(ct$source, levels = src_order)
  }
  if (!is.null(tgt_order)) {
    ct$target <- factor(ct$target, levels = tgt_order)
  }
  ct$source <- as.character(ct$source)
  ct$target <- as.character(ct$target)
  ct
}

make_policy_tables <- function(obs_df) {
  by_year <- obs_df %>% group_by(Year) %>%
    summarise(n = n(), confirmed = sum(HAS_CONFIRMED_PLE, na.rm=TRUE),
              no_match = n() - confirmed,
              pct = round(confirmed/n*100, 2), .groups="drop") %>% arrange(Year)
  by_course <- obs_df %>% group_by(CourseGroup) %>%
    summarise(n = n(), confirmed = sum(HAS_CONFIRMED_PLE, na.rm=TRUE),
              no_match = n() - confirmed,
              pct = round(confirmed/n*100, 2),
              med_pct = round(median(NMS_PER_num, na.rm=TRUE), 2), .groups="drop") %>%
    arrange(desc(pct))
  by_uni <- obs_df %>% filter(UNI_TYPE %in% c("Public","Private","Foreign")) %>%
    group_by(UNI_TYPE) %>%
    summarise(n = n(), confirmed = sum(HAS_CONFIRMED_PLE, na.rm=TRUE),
              no_match = n() - confirmed,
              pct = round(confirmed/n*100, 2),
              med_pct = round(median(NMS_PER_num, na.rm=TRUE), 2), .groups="drop")
  list(by_year = by_year, by_course = by_course, by_uni = by_uni)
}

bin_heatmap <- function(df, group_col, title) {
  res <- pct_cross(df, group_col, "PercentileBin", BIN_ORDER)
  mat <- as.matrix(res$pct[, BIN_ORDER])
  rownames(mat) <- res$pct[[group_col]]
  plot_ly(z = mat, x = BIN_ORDER, y = rownames(mat), type = "heatmap",
          colorscale = "YlOrRd", text = round(mat, 1),
          texttemplate = "%{text}", showscale = TRUE) %>%
    layout(title = title, xaxis = list(title = "Bin"),
           yaxis = list(title = ""))
}

# ── PRE-COMPUTE DERIVED DATASETS ──────────────────────────
F_besttrend <- DF_BESTTREND
F_bestobs   <- DF_BESTOBS
F_uni       <- DF_UNI
F_uniobs    <- DF_UNIOBS
F_trend     <- DF_TREND

gap_df <- F_bestobs %>%
  filter(PLE_STATUS_LABEL == "Confirmed PLE passer", !is.na(PLE_YEAR_GAP))

sex_base <- F_besttrend %>% filter(!is.na(SEX_CLEAN))

repeat_data <- local({
  df <- F_trend
  attempt <- df %>% group_by(PERSON_KEY) %>%
    summarise(n_attempts = n_distinct(APPNO_CLEAN), .groups="drop")
  repeaters <- attempt %>% filter(n_attempts > 1) %>% pull(PERSON_KEY)
  rep_df <- df %>% filter(PERSON_KEY %in% repeaters) %>%
    arrange(PERSON_KEY, Year, APPNO_CLEAN) %>%
    group_by(PERSON_KEY) %>%
    summarise(
      first_year = first(Year), last_year = last(Year),
      first_pct  = first(NMS_PER_num), last_pct  = last(NMS_PER_num),
      first_raw  = first(TotalRawScoreTRUE), last_raw = last(TotalRawScoreTRUE),
      n_attempts = n_distinct(APPNO_CLEAN), .groups="drop") %>%
    filter(!is.na(first_pct), !is.na(last_pct)) %>%
    mutate(pct_change = last_pct - first_pct, raw_change = last_raw - first_raw)
  list(attempt = attempt, rep_df = rep_df)
})

cmp_df <- local({
  parts <- list()
  
  g1 <- F_uniobs[!is.na(F_uniobs$FOREIGNER_STATUS) & F_uniobs$FOREIGNER_STATUS == "Verified Foreigner", ]
  if (nrow(g1) > 0) { g1$cmp_group <- "Foreigners (non-Filipino)"; parts[[1]] <- g1 }
  
  g2 <- F_uniobs[!is.na(F_uniobs$CITIZENSHIP_FINAL) & F_uniobs$CITIZENSHIP_FINAL == "Filipino" & !is.na(F_uniobs$UNI_TYPE) & F_uniobs$UNI_TYPE == "Foreign", ]
  if (nrow(g2) > 0) { g2$cmp_group <- "Filipinos at Foreign Schools"; parts[[2]] <- g2 }
  
  g3 <- F_uniobs[!is.na(F_uniobs$UNI_TYPE) & F_uniobs$UNI_TYPE == "Public", ]
  if (nrow(g3) > 0) { g3$cmp_group <- "Filipinos at Public Schools"; parts[[3]] <- g3 }
  
  g4 <- F_uniobs[!is.na(F_uniobs$UNI_TYPE) & F_uniobs$UNI_TYPE == "Private", ]
  if (nrow(g4) > 0) { g4$cmp_group <- "Filipinos at Private Schools"; parts[[4]] <- g4 }
  
  if (length(parts) == 0) return(NULL)
  shared <- c("cmp_group","NMS_PER_num","TotalRawScoreTRUE","PercentileBin",
              "HAS_CONFIRMED_PLE","UNI_TYPE","CourseGroup","Year")
  dplyr::bind_rows(lapply(parts, function(p) p[, intersect(shared, names(p)), drop = FALSE]))
})

policy_tbls <- make_policy_tables(F_bestobs)

survival_tbl <- F_besttrend %>%
  filter(!is.na(CourseGroup), !is.na(PercentileBin)) %>%
  group_by(CourseGroup) %>%
  summarise(
    total  = n(),
    top_n  = sum(PercentileBin %in% c("B8","B9","B10"), na.rm = TRUE),
    top_pct = round(top_n / total * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(top_pct))
```

# Executive Summary

<div class="callout-info">
<strong>Overview:</strong> Key performance metrics, score distributions, and institutional composition
for NMAT examinees from 2006 to 2018. All figures use the best NMAT record per person and the
TRUE raw score where available. The observable cohort (NMAT year &le; 2014) is used for PLE-linked analyses
to avoid misclassifying later cohorts as non-passers before their licensure window closes.
</div>

```{r p01_metrics, echo=FALSE, results="asis"}
n_best    <- nrow(F_besttrend)
n_years   <- dplyr::n_distinct(F_besttrend$Year)
med_raw   <- round(median(F_besttrend$TotalRawScoreTRUE, na.rm=TRUE), 1)
med_pct   <- round(median(F_besttrend$NMS_PER_num, na.rm=TRUE), 1)
n_unique  <- dplyr::n_distinct(F_besttrend$PERSON_KEY)
n_repeat  <- local({
  ct <- F_trend %>% dplyr::group_by(PERSON_KEY) %>%
    dplyr::summarise(n = dplyr::n_distinct(APPNO_CLEAN), .groups="drop")
  sum(ct$n > 1)
})
n_obs     <- nrow(F_bestobs)
ple_share <- round(mean(F_bestobs$HAS_CONFIRMED_PLE, na.rm=TRUE) * 100, 2)

card <- function(value, label) {
  paste0(
    '<div class="metric-card">',
    '<div class="metric-value">', value, '</div>',
    '<div class="metric-label">', label, '</div>',
    '</div>'
  )
}

cat(paste0(
  '<div class="metric-row">',
  card(format(n_best,   big.mark=","), "Best-record examinees"),
  card(n_years,                        "Years covered (2006–2018)"),
  card(med_raw,                        "Median TRUE raw score"),
  card(med_pct,                        "Median percentile rank"),
  card(format(n_unique, big.mark=","), "Unique examinees"),
  card(format(n_repeat, big.mark=","), "Repeat takers"),
  card(format(n_obs,    big.mark=","), "Observable cohort size"),
  card(paste0(ple_share, "%"),         "Confirmed PLE share (observable)"),
  '</div>'
))
```

```{r p01_trend, echo=FALSE, fig.height=9}
s <- yearly_summary(F_besttrend)
fig_p01 <- subplot(
  plot_ly(s, x = ~Year) %>%
    add_ribbons(ymin = ~raw_q25, ymax = ~raw_q75,
                fillcolor = "rgba(31,119,180,.25)", line = list(color = "lightblue"),
                name = "IQR raw", showlegend = FALSE) %>%
    add_lines(y = ~raw_median, name = "Median raw",
              line = list(color = "#1f77b4", width = 3)) %>%
    layout(yaxis = list(title = "Total Raw Score")),
  plot_ly(s, x = ~Year) %>%
    add_lines(y = ~part1_median, name = "Part I",
              line = list(color = "#2ca02c", width = 3)) %>%
    add_lines(y = ~part2_median, name = "Part II",
              line = list(color = "#ff7f0e", width = 3)) %>%
    layout(yaxis = list(title = "Part Score")),
  plot_ly(s, x = ~Year) %>%
    add_ribbons(ymin = ~per_q25, ymax = ~per_q75,
                fillcolor = "rgba(255,127,14,.22)", line = list(color = "lightsalmon"),
                name = "IQR pct", showlegend = FALSE) %>%
    add_lines(y = ~per_median, name = "Median pct",
              line = list(color = "#d62728", width = 3)) %>%
    layout(yaxis = list(title = "Percentile Rank")),
  plot_ly(s, x = ~Year, y = ~n, type = "bar", name = "Examinees") %>%
    layout(yaxis = list(title = "Count")),
  nrows = 2, shareX = FALSE, titleX = TRUE
) %>% layout(
  height = 680,
  hovermode = "x unified",
  annotations = list(
    list(text = "Median Raw Score (±IQR)", x = 0.225, y = 1.03,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333")),
    list(text = "Part I vs Part II Raw Score", x = 0.775, y = 1.03,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333")),
    list(text = "Median Percentile (±IQR)", x = 0.225, y = 0.47,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333")),
    list(text = "Examinee Count by Year", x = 0.775, y = 0.47,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333"))
  )
)
fig_p01
```

```{r p01_trend_tbl, echo=FALSE}
dt_export(s, "Annual NMAT Score Summary by Year")
```

```{r p01_course_pie, echo=FALSE, fig.height=5}
ct_course <- F_besttrend %>% dplyr::count(CourseGroup)
plot_ly(ct_course, labels = ~CourseGroup, values = ~n, type = "pie",
        marker = list(colors = PAL_COURSE[ct_course$CourseGroup])) %>%
  layout(title = "Course-Group Composition")
```

```{r p01_course_tbl, echo=FALSE}
dt_export(ct_course, "Course Group Composition - Counts")
```

```{r p01_uni_pie, echo=FALSE, fig.height=5}
ct_uni <- F_besttrend %>% dplyr::count(UNI_TYPE)
plot_ly(ct_uni, labels = ~UNI_TYPE, values = ~n, type = "pie",
        marker = list(colors = PAL_UNI[ct_uni$UNI_TYPE])) %>%
  layout(title = "University-Type Composition")
```

```{r p01_uni_tbl, echo=FALSE}
dt_export(ct_uni, "University Type Composition - Counts")
```

```{r p01_summary_tbl, echo=FALSE}
df_p01 <- F_besttrend
tbl_p01 <- tibble::tibble(
  Indicator = c("Median Total Raw Score","Median Part I Raw Score",
                "Median Part II Raw Score","Median Percentile Rank",
                "Top-bin share B8-B10 (%)","Bottom-bin share B1-B3 (%)"),
  Value = c(
    round(median(df_p01$TotalRawScoreTRUE, na.rm=TRUE), 1),
    round(median(df_p01$PartIRawScoreTRUE, na.rm=TRUE), 1),
    round(median(df_p01$PartIIRawScoreTRUE, na.rm=TRUE), 1),
    round(median(df_p01$NMS_PER_num, na.rm=TRUE), 1),
    round(mean(df_p01$PercentileBin %in% c("B8","B9","B10"), na.rm=TRUE)*100, 2),
    round(mean(df_p01$PercentileBin %in% c("B1","B2","B3"), na.rm=TRUE)*100, 2)
  )
)
dt_export(tbl_p01, "Summary Indicators")
```

# Data Integrity

<div class="callout-info">
<strong>Cohort definitions:</strong> All rows = every cleaned NMAT record. Best-record = highest percentile per person.
Observable cohort = best-record examinees with NMAT year &le; 2014 (enough time has passed to observe a PLE outcome).
Raw-score validation checks the Part I + Part II = Total formula.
</div>

```{r p02_metrics, echo=FALSE, results="asis"}
n_all_rows  <- nrow(DF_RAW)
n_best_rows <- nrow(DF_BEST)
n_true_raw  <- sum(DF_RAW$HasTRUErawScores == TRUE, na.rm=TRUE)
n_obs_rows  <- nrow(DF_BESTOBS)

cat(paste0(
  '<div class="metric-row">',
  card(format(n_all_rows,  big.mark=","), "All NMAT rows"),
  card(format(n_best_rows, big.mark=","), "Best-record rows"),
  card(format(n_true_raw,  big.mark=","), "Rows with TRUE raw scores"),
  card(format(n_obs_rows,  big.mark=","), "Observable best-record rows"),
  '</div>'
))
```

## Cohort Overview

```{r p02_cohort_tbl, echo=FALSE}
tbl_cohort <- tibble::tibble(
  `Analytic subset` = c("All cleaned NMAT rows","Best record per person",
                        "Best-record 2006-2018","Best-record (observable, <=2014)",
                        "Confirmed PLE-matched rows"),
  `Row count` = c(nrow(DF_RAW), nrow(DF_BEST), nrow(DF_BESTTREND),
                  nrow(DF_BESTOBS),
                  sum(DF_RAW$IS_PLE_ANALYSIS_SAFE == TRUE, na.rm=TRUE))
)
dt_export(tbl_cohort, "Analysis Cohorts")
```

## Raw Score Validation

```{r p02_rawscore_tbl, echo=FALSE}
df_p02 <- DF_RAW
ok_p02  <- complete.cases(df_p02[, c("TotalRawScoreTRUE","PartIRawScoreTRUE","PartIIRawScoreTRUE")])
mis_p02 <- abs(df_p02$TotalRawScoreTRUE[ok_p02] -
               df_p02$PartIRawScoreTRUE[ok_p02] -
               df_p02$PartIIRawScoreTRUE[ok_p02]) > 0.01
svd_p02 <- if ("StoredVsDerivedMismatch" %in% names(df_p02))
             sum(as.numeric(df_p02$StoredVsDerivedMismatch) > 0, na.rm=TRUE) else NA_integer_
cvd_p02 <- if ("CalcVsDerivedMismatch" %in% names(df_p02))
             sum(as.numeric(df_p02$CalcVsDerivedMismatch) > 0, na.rm=TRUE) else NA_integer_
tbl_rawscore <- tibble::tibble(
  Check = c("Complete Total+Part I+Part II rows",
            "Formula mismatches (Total != I + II)",
            "Stored-vs-derived mismatch flag",
            "Calc-vs-derived mismatch flag"),
  Count = c(sum(ok_p02), sum(mis_p02, na.rm=TRUE), svd_p02, cvd_p02)
)
dt_export(tbl_rawscore, "Raw-Score Validation")
```

## Distributions

```{r p02_unitype_dist, echo=FALSE}
dt_export(as.data.frame(table(UNI_TYPE = DF_RAW$UNI_TYPE)), "UNI_TYPE Distribution")
```

```{r p02_course_dist, echo=FALSE}
dt_export(as.data.frame(table(CourseGroup = DF_RAW$CourseGroup)), "Course Group Distribution")
```

```{r p02_ple_dist, echo=FALSE}
dt_export(as.data.frame(table(PLE_Status = as.character(DF_RAW$PLE_STATUS_LABEL))), "PLE Status Distribution")
```

# Performance Trends

<div class="callout-info">
<strong>What to look for:</strong> The four-panel trend chart shows median raw scores, part scores, percentile ranks, and
exam volume across 2006&ndash;2018. Shaded bands show the interquartile range (IQR).
The Kruskal-Wallis table tests whether score distributions differ significantly across years;
eta-squared &ge; 0.01 indicates a practically meaningful effect.
</div>

## Trend Overview

```{r p03_trend_fig, echo=FALSE, fig.height=9}
s_p03 <- yearly_summary(F_besttrend)
fig_p03 <- subplot(
  plot_ly(s_p03, x = ~Year) %>%
    add_ribbons(ymin = ~raw_q25, ymax = ~raw_q75,
                fillcolor = "rgba(31,119,180,.2)", line = list(color="lightblue"),
                showlegend = FALSE) %>%
    add_lines(y = ~raw_median, line = list(color="#1f77b4",width=3), name="Median raw"),
  plot_ly(s_p03, x = ~Year) %>%
    add_lines(y = ~part1_median, name="Part I",  line=list(color="#2ca02c",width=3)) %>%
    add_lines(y = ~part2_median, name="Part II", line=list(color="#ff7f0e",width=3)),
  plot_ly(s_p03, x = ~Year) %>%
    add_ribbons(ymin = ~per_q25, ymax = ~per_q75,
                fillcolor = "rgba(214,39,40,.15)", line=list(color="lightsalmon"),
                showlegend=FALSE) %>%
    add_lines(y = ~per_median, line=list(color="#d62728",width=3), name="Median pct"),
  plot_ly(s_p03, x = ~Year, y = ~n, type="bar", name="Examinees"),
  nrows=2, shareX=FALSE, titleX=TRUE
) %>% layout(
  height = 700,
  hovermode = "x unified",
  annotations = list(
    list(text = "Median Raw Score (±IQR)", x = 0.225, y = 1.03,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333")),
    list(text = "Part I vs Part II Raw Score", x = 0.775, y = 1.03,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333")),
    list(text = "Median Percentile (±IQR)", x = 0.225, y = 0.47,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333")),
    list(text = "Examinee Count by Year", x = 0.775, y = 0.47,
         xref = "paper", yref = "paper", showarrow = FALSE,
         font = list(size = 13, color = "#333"))
  )
)
fig_p03
```

```{r p03_trend_tbl, echo=FALSE}
dt_export(s_p03, "Annual Score Summary (Performance Trends)")
```

## Score Distributions by Year

```{r p03_box_raw, echo=FALSE, fig.height=5}
df_p03_raw <- F_besttrend %>% filter(!is.na(Year), !is.na(TotalRawScoreTRUE))
plot_ly(df_p03_raw, x = ~factor(Year), y = ~TotalRawScoreTRUE, type="box", boxpoints=FALSE,
        marker=list(color="#1f77b4")) %>%
  layout(title="Total Raw Score by Year", xaxis=list(title="Year"),
         yaxis=list(title="Score"))
```

```{r p03_raw_yr_tbl, echo=FALSE}
tbl_p03_raw_yr <- F_besttrend %>%
  dplyr::filter(!is.na(Year), !is.na(TotalRawScoreTRUE)) %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(n = dplyr::n(),
    median = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
    q25    = round(quantile(TotalRawScoreTRUE, .25, na.rm = TRUE), 2),
    q75    = round(quantile(TotalRawScoreTRUE, .75, na.rm = TRUE), 2),
    .groups = "drop")
dt_export(tbl_p03_raw_yr, "Total Raw Score Statistics by Year")
```

```{r p03_box_pct, echo=FALSE, fig.height=5}
df_p03_pct <- F_besttrend %>% filter(!is.na(Year), !is.na(NMS_PER_num))
plot_ly(df_p03_pct, x = ~factor(Year), y = ~NMS_PER_num, type="box", boxpoints=FALSE,
        marker=list(color="#d62728")) %>%
  layout(title="Percentile Rank by Year", xaxis=list(title="Year"),
         yaxis=list(title="Percentile"))
```

```{r p03_pct_yr_tbl, echo=FALSE}
tbl_p03_pct_yr <- F_besttrend %>%
  dplyr::filter(!is.na(Year), !is.na(NMS_PER_num)) %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(n = dplyr::n(),
    median = round(median(NMS_PER_num, na.rm = TRUE), 2),
    q25    = round(quantile(NMS_PER_num, .25, na.rm = TRUE), 2),
    q75    = round(quantile(NMS_PER_num, .75, na.rm = TRUE), 2),
    .groups = "drop")
dt_export(tbl_p03_pct_yr, "Percentile Rank Statistics by Year")
```

## Statistical Tests

```{r p03_kw_tbl, echo=FALSE}
kw_res <- kw_table(F_besttrend, "Year",
  list(TotalRawScoreTRUE="Total Raw Score", PartIRawScoreTRUE="Part I",
       PartIIRawScoreTRUE="Part II", NMS_PER_num="Percentile Rank",
       NMS_GPS="GPS Standard Score"))
dt_export(kw_res, "Kruskal-Wallis — Score Distributions Across Years")
```

# Score Bins & Background

## By Year

```{r p04_heatmap_year}
# Bin % heatmap by Year (YlOrRd)
res_yr <- pct_cross(F_besttrend, "Year", "PercentileBin", BIN_ORDER)
mat_yr <- as.matrix(res_yr$pct[, BIN_ORDER])
rownames(mat_yr) <- res_yr$pct[["Year"]]
plot_ly(
  z             = mat_yr,
  x             = BIN_ORDER,
  y             = rownames(mat_yr),
  type          = "heatmap",
  colorscale    = "YlOrRd",
  text          = round(mat_yr, 1),
  texttemplate  = "%{text}",
  showscale     = TRUE
) %>%
  layout(
    title  = "Bin % by Year",
    xaxis  = list(title = "Bin"),
    yaxis  = list(title = "")
  )
```

```{r p04_heatmap_year_pct, echo=FALSE}
dt_export(res_yr$pct %>% dplyr::mutate(Year = as.character(Year)),
          "Bin Percentages by Year (row %)")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Each cell shows the percentage of examinees in that bin for the given year. Darker = higher concentration.</p>

---

```{r p04_count_year}
# Raw count table by Year × Bin
res_yr_ct <- pct_cross(F_besttrend, "Year", "PercentileBin", BIN_ORDER)
dt_export(res_yr_ct$counts, "Bin Counts by Year")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Raw count of best-record examinees per bin per year.</p>

---

```{r p04_stacked_year}
# Stacked bar: bin composition % by exam year
res_stk <- pct_cross(F_besttrend, "Year", "PercentileBin", BIN_ORDER)
long_stk <- res_stk$pct %>%
  dplyr::mutate(Year = as.integer(Year)) %>%
  tidyr::pivot_longer(-Year, names_to = "Bin", values_to = "Pct") %>%
  dplyr::mutate(Bin = factor(Bin, BIN_ORDER))
plot_ly(long_stk, x = ~Year, y = ~Pct, color = ~Bin, type = "bar",
        colors = PAL_BIN) %>%
  layout(barmode = "stack",
         title   = "Bin Composition by Year",
         yaxis   = list(title = "Percent"),
         xaxis   = list(title = "Year"))
```

```{r p04_stacked_yr_tbl, echo=FALSE}
dt_export(res_stk$pct %>% dplyr::mutate(Year = as.character(Year)),
          "Bin Composition by Year (%)")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Stacked bars: bin composition (%) by exam year.</p>

---

```{r p04_topbottom_year}
# Top/bottom trend lines: B8-B10 vs B1-B3 over time
res_tb <- pct_cross(F_besttrend, "Year", "PercentileBin", BIN_ORDER)
s_tb <- res_tb$pct %>%
  dplyr::mutate(
    Year   = as.integer(Year),
    top    = rowSums(dplyr::select(., dplyr::any_of(c("B8","B9","B10"))), na.rm = TRUE),
    bottom = rowSums(dplyr::select(., dplyr::any_of(c("B1","B2","B3"))),  na.rm = TRUE)
  )
plot_ly(s_tb, x = ~Year) %>%
  add_lines(y = ~top,    name = "Top B8-B10",   line = list(color = "#2e7d32", width = 3)) %>%
  add_lines(y = ~bottom, name = "Bottom B1-B3", line = list(color = "#c62828", width = 3)) %>%
  layout(
    title  = "Top vs Bottom Bin Share by Year",
    yaxis  = list(title = "Percent"),
    xaxis  = list(title = "Year")
  )
```

```{r p04_topbot_tbl, echo=FALSE}
tbl_tb <- s_tb %>%
  dplyr::select(Year, top, bottom) %>%
  dplyr::rename(Top_B8_B10_pct = top, Bottom_B1_B3_pct = bottom)
dt_export(tbl_tb, "Top vs Bottom Bin Share by Year")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Trend lines: share of examinees in the top-scoring (B8-B10) vs bottom-scoring (B1-B3) bins over time.</p>

## By University Type

```{r p04_heatmap_uni}
# Bin % heatmap by UNI_TYPE (YlOrRd)
res_uni <- pct_cross(F_uni, "UNI_TYPE", "PercentileBin", BIN_ORDER)
mat_uni <- as.matrix(res_uni$pct[, BIN_ORDER])
rownames(mat_uni) <- res_uni$pct[["UNI_TYPE"]]
plot_ly(
  z             = mat_uni,
  x             = BIN_ORDER,
  y             = rownames(mat_uni),
  type          = "heatmap",
  colorscale    = "YlOrRd",
  text          = round(mat_uni, 1),
  texttemplate  = "%{text}",
  showscale     = TRUE
) %>%
  layout(
    title  = "Bin % by University Type",
    xaxis  = list(title = "Bin"),
    yaxis  = list(title = "")
  )
```

```{r p04_heatmap_uni_tbl, echo=FALSE}
dt_export(res_uni$pct, "Bin Percentages by University Type (row %)")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Row percentages: each cell shows the % of examinees from that university type in each bin.</p>

---

```{r p04_top_uni}
# Top-bin bar: B8-B10 % by UNI_TYPE
res_tu <- pct_cross(F_uni, "UNI_TYPE", "PercentileBin", BIN_ORDER)
top_uni <- res_tu$pct %>%
  dplyr::mutate(Top_B8_B10 = B8 + B9 + B10) %>%
  dplyr::select(UNI_TYPE, Top_B8_B10) %>%
  dplyr::arrange(Top_B8_B10)
plot_ly(top_uni,
        y           = ~UNI_TYPE,
        x           = ~Top_B8_B10,
        type        = "bar",
        orientation = "h",
        marker      = list(color = PAL_UNI[top_uni$UNI_TYPE])) %>%
  layout(
    title  = "Top Bin Share by University Type",
    xaxis  = list(title = "Percent in B8-B10"),
    yaxis  = list(title = "")
  )
```

```{r p04_top_uni_tbl, echo=FALSE}
dt_export(top_uni, "Top Bin Share (B8-B10) by University Type")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Share of examinees from each university type that scored in the top bins (B8-B10).</p>

---

```{r p04_chi}
# Chi-square summary table
df_chi  <- F_uni %>% dplyr::filter(!is.na(UNI_TYPE), !is.na(PercentileBin))
ct_chi  <- table(df_chi$UNI_TYPE, df_chi$PercentileBin)
tryCatch({
  ch_res  <- chisq.test(ct_chi)
  n_chi   <- sum(ct_chi)
  r_chi   <- nrow(ct_chi)
  c_chi   <- ncol(ct_chi)
  V_chi   <- sqrt(ch_res$statistic / (n_chi * min(r_chi - 1, c_chi - 1)))
  tbl_chi <- tibble::tibble(
    chi2      = round(ch_res$statistic, 4),
    p_value   = ifelse(ch_res$p.value < .001, "<0.001", round(ch_res$p.value, 4)),
    df        = ch_res$parameter,
    n         = n_chi,
    cramers_v = round(V_chi, 4)
  )
  dt_export(tbl_chi, "Chi-Square: University Type x Bin")
}, error = function(e) {
  dt_export(tibble::tibble(Error = conditionMessage(e)))
})
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Chi-square test for independence between university type and percentile bin. Cramer's V measures effect size (0 = no association, 1 = perfect).</p>

---

<div style="background:#f0f4ff;border-left:3px solid #1a237e;padding:8px 14px;margin:18px 0 10px 0;border-radius:2px;font-size:14px;font-weight:600;color:#1a237e;">Bin distribution by year, faceted by university type</div>
<p style="color:#555;font-size:13px;margin-bottom:8px;">Each panel shows one university type. Bars are stacked by percentile bin. Use this to track how the bin composition of each sector evolved across exam years.</p>

```{r p04_faceted}
# Faceted stacked bar: 3 panels Foreign/Private/Public, bins by year
df_fac <- F_uni %>%
  dplyr::filter(UNI_TYPE %in% c("Public","Private","Foreign"),
                !is.na(PercentileBin), !is.na(Year)) %>%
  dplyr::count(Year, UNI_TYPE, PercentileBin) %>%
  dplyr::mutate(
    PercentileBin = factor(PercentileBin, BIN_ORDER),
    Year          = as.character(Year)
  )

if (nrow(df_fac) == 0) {
  plot_ly() %>% layout(title = "No data available")
} else {
  fig_fac  <- plot_ly()
  uni_types_fac <- c("Foreign","Private","Public")
  for (i in seq_along(uni_types_fac)) {
    sub_fac <- df_fac %>% dplyr::filter(UNI_TYPE == uni_types_fac[i])
    for (b in BIN_ORDER) {
      bsub_fac <- sub_fac %>% dplyr::filter(PercentileBin == b)
      if (nrow(bsub_fac) == 0) next
      fig_fac <- fig_fac %>% add_bars(
        data        = bsub_fac,
        x           = ~Year,
        y           = ~n,
        name        = b,
        legendgroup = b,
        showlegend  = (i == 1),
        marker      = list(color = PAL_BIN[[b]]),
        xaxis       = if (i == 1) "x" else paste0("x", i),
        yaxis       = if (i == 1) "y" else paste0("y", i)
      )
    }
  }
  fig_fac %>% layout(
    barmode = "stack",
    title   = "Examinees per bin by year and university type",
    grid    = list(rows = 1, columns = 3, pattern = "independent"),
    xaxis   = list(title = "Foreign",  domain = c(0,    0.30)),
    xaxis2  = list(title = "Private",  domain = c(0.35, 0.65)),
    xaxis3  = list(title = "Public",   domain = c(0.70, 1.00)),
    yaxis   = list(title = "Examinees"),
    yaxis2  = list(title = ""),
    yaxis3  = list(title = ""),
    height  = 520,
    annotations = list(
      list(x = 0.15, y = 1.05, xref = "paper", yref = "paper",
           text = "Foreign",  showarrow = FALSE, font = list(size = 13)),
      list(x = 0.50, y = 1.05, xref = "paper", yref = "paper",
           text = "Private",  showarrow = FALSE, font = list(size = 13)),
      list(x = 0.85, y = 1.05, xref = "paper", yref = "paper",
           text = "Public",   showarrow = FALSE, font = list(size = 13))
    )
  )
}
```

```{r p04_faceted_tbl, echo=FALSE}
df_fac_wide <- df_fac %>%
  tidyr::pivot_wider(names_from = PercentileBin, values_from = n, values_fill = 0L)
dt_export(df_fac_wide, "Examinees by Year, University Type, and Bin (Counts)")
```

---

<div style="background:#f0f4ff;border-left:3px solid #1a237e;padding:8px 14px;margin:18px 0 10px 0;border-radius:2px;font-size:14px;font-weight:600;color:#1a237e;">Record-level listings by university type</div>
<p style="color:#555;font-size:13px;margin-bottom:8px;">All best-record examinees (2006-2018) for each university type, sorted by year (newest first) then percentile rank (highest first).</p>

#### Foreign Records

```{r p04_rec_foreign}
rec_cols_p04 <- intersect(
  c("PERSON_KEY","APPNO_CLEAN","UNIVERSITY","UNI_LOCATION","Year","CourseGroup",
    "PercentileBin","NMS_PER_num","TotalRawScoreTRUE"),
  names(F_uni)
)
sub_foreign <- F_uni %>%
  dplyr::filter(UNI_TYPE == "Foreign") %>%
  dplyr::arrange(dplyr::desc(Year), dplyr::desc(NMS_PER_num)) %>%
  dplyr::select(dplyr::all_of(rec_cols_p04))
dt_export(sub_foreign, "Foreign examinee records")
```

#### Public Records

```{r p04_rec_public}
sub_public <- F_uni %>%
  dplyr::filter(UNI_TYPE == "Public") %>%
  dplyr::arrange(dplyr::desc(Year), dplyr::desc(NMS_PER_num)) %>%
  dplyr::select(dplyr::all_of(rec_cols_p04))
dt_export(sub_public, "Public university examinee records")
```

#### Private Records

```{r p04_rec_private}
sub_private <- F_uni %>%
  dplyr::filter(UNI_TYPE == "Private") %>%
  dplyr::arrange(dplyr::desc(Year), dplyr::desc(NMS_PER_num)) %>%
  dplyr::select(dplyr::all_of(rec_cols_p04))
dt_export(sub_private, "Private university examinee records")
```

---

<div style="background:#f0f4ff;border-left:3px solid #1a237e;padding:8px 14px;margin:18px 0 10px 0;border-radius:2px;font-size:14px;font-weight:600;color:#1a237e;">No-PLE-match listings by university type</div>
<div style="background:#fff8e1;border:1px solid #ffe082;padding:8px 12px;border-radius:3px;font-size:13px;margin:8px 0;"><b>Observable cohort only (NMAT year &le; 2014):</b> Examinees who sat the NMAT in 2014 or earlier but have no confirmed PLE match. This cohort has had sufficient time to attempt the boards; the absence of a match is analytically meaningful rather than a data lag.</div>

```{r p04_nople_cols_setup, echo=FALSE}
nople_cols_p04 <- intersect(
  c("PERSON_KEY","APPNO_CLEAN","UNIVERSITY","UNI_LOCATION","Year","CourseGroup",
    "PercentileBin","NMS_PER_num","TotalRawScoreTRUE","PLE_STATUS_LABEL"),
  names(F_uniobs)
)
```

#### Foreign No-PLE

```{r p04_nople_foreign}
nople_foreign <- F_uniobs %>%
  dplyr::filter(UNI_TYPE == "Foreign",
                PLE_STATUS_LABEL == "No confirmed PLE match") %>%
  dplyr::arrange(dplyr::desc(Year), dplyr::desc(NMS_PER_num)) %>%
  dplyr::select(dplyr::all_of(nople_cols_p04))
dt_export(nople_foreign, "Foreign - No PLE match (observable cohort)")
```

#### Public No-PLE

```{r p04_nople_public}
nople_public <- F_uniobs %>%
  dplyr::filter(UNI_TYPE == "Public",
                PLE_STATUS_LABEL == "No confirmed PLE match") %>%
  dplyr::arrange(dplyr::desc(Year), dplyr::desc(NMS_PER_num)) %>%
  dplyr::select(dplyr::all_of(nople_cols_p04))
dt_export(nople_public, "Public - No PLE match (observable cohort)")
```

#### Private No-PLE

```{r p04_nople_private}
nople_private <- F_uniobs %>%
  dplyr::filter(UNI_TYPE == "Private",
                PLE_STATUS_LABEL == "No confirmed PLE match") %>%
  dplyr::arrange(dplyr::desc(Year), dplyr::desc(NMS_PER_num)) %>%
  dplyr::select(dplyr::all_of(nople_cols_p04))
dt_export(nople_private, "Private - No PLE match (observable cohort)")
```

---

<div style="background:#f0f4ff;border-left:3px solid #1a237e;padding:8px 14px;margin:18px 0 10px 0;border-radius:2px;font-size:14px;font-weight:600;color:#1a237e;">Citizenship profile - no-PLE-match examinees</div>
<p style="color:#555;font-size:13px;margin-bottom:8px;">Links no-PLE-match records to a profiling file via APPNOCLEAN. Citizenship is an inferred nationality label derived from name patterns and registration data. The metrics below describe NMAT performance by inferred citizenship group.</p>

#### Citizenship Profile

```{r p04_pseudo_setup, echo=FALSE}
pc_base <- F_uniobs[!is.na(F_uniobs$PLE_STATUS_LABEL) & F_uniobs$PLE_STATUS_LABEL == "No confirmed PLE match" & !is.na(F_uniobs$CITIZENSHIP_FINAL), ]
pc_available <- nrow(pc_base) > 0
```

```{r p04_pc_metrics, results="asis", echo=FALSE}
if (pc_available) {
  n_total    <- nrow(pc_base)
  n_foreign  <- if ("FOREIGNER_STATUS" %in% names(pc_base))
                  sum(pc_base$FOREIGNER_STATUS == "Verified Foreigner", na.rm = TRUE) else 0L
  n_filipino <- sum(pc_base$CITIZENSHIP_FINAL == "Filipino", na.rm = TRUE)
  n_distinct_labels <- dplyr::n_distinct(pc_base$CITIZENSHIP_FINAL)
  cat(sprintf(
    '<div style="display:flex;gap:16px;flex-wrap:wrap;margin-bottom:16px;">
      <div style="background:#e8eaf6;border-radius:6px;padding:12px 20px;min-width:160px;">
        <div style="font-size:24px;font-weight:700;color:#3949ab;">%s</div>
        <div style="font-size:12px;color:#555;">Profiled no-PLE-match records</div>
      </div>
      <div style="background:#f3e5f5;border-radius:6px;padding:12px 20px;min-width:160px;">
        <div style="font-size:24px;font-weight:700;color:#7b1fa2;">%s</div>
        <div style="font-size:12px;color:#555;">Foreigners</div>
      </div>
      <div style="background:#e3f2fd;border-radius:6px;padding:12px 20px;min-width:160px;">
        <div style="font-size:24px;font-weight:700;color:#1565c0;">%s</div>
        <div style="font-size:12px;color:#555;">Filipinos</div>
      </div>
      <div style="background:#e8f5e9;border-radius:6px;padding:12px 20px;min-width:160px;">
        <div style="font-size:24px;font-weight:700;color:#2e7d32;">%s</div>
        <div style="font-size:12px;color:#555;">Distinct citizenship labels</div>
      </div>
    </div>',
    format(n_total, big.mark = ","),
    format(n_foreign, big.mark = ","),
    format(n_filipino, big.mark = ","),
    format(n_distinct_labels, big.mark = ",")
  ))
}
```

```{r p04_pc_pie_comp}
if (pc_available) {
  ct_pc <- pc_base %>% dplyr::count(CITIZENSHIP_FINAL) %>% dplyr::arrange(dplyr::desc(n))
  plot_ly(ct_pc, labels = ~CITIZENSHIP_FINAL, values = ~n, type = "pie") %>%
    layout(title = "Citizenship composition", showlegend = TRUE)
}
```

```{r p04_pc_pie_fvf}
if (pc_available) {
  nf_pc <- if ("FOREIGNER_STATUS" %in% names(pc_base))
    sum(pc_base$FOREIGNER_STATUS == "Verified Foreigner", na.rm = TRUE) else 0L
  np_pc <- sum(pc_base$CITIZENSHIP_FINAL == "Filipino", na.rm = TRUE)
  ct_fvf <- data.frame(group = c("Foreigner","Filipino"), n = c(nf_pc, np_pc))
  plot_ly(ct_fvf, labels = ~group, values = ~n, type = "pie",
          marker = list(colors = c("#9467bd","#1f77b4"))) %>%
    layout(title = "Foreigners vs Filipinos")
}
```

```{r p04_pc_fvf_tbl, echo=FALSE}
if (pc_available) dt_export(ct_fvf, "Foreigners vs Filipinos (Counts)")
```

```{r p04_pc_top15}
if (pc_available) {
  top15_pc <- pc_base %>%
    dplyr::count(CITIZENSHIP_FINAL) %>%
    dplyr::arrange(dplyr::desc(n)) %>%
    head(15)
  plot_ly(top15_pc,
          x            = ~n,
          y            = ~reorder(CITIZENSHIP_FINAL, n),
          type         = "bar",
          orientation  = "h",
          marker       = list(color = "#9467bd"),
          text         = ~as.character(n),
          textposition = "outside") %>%
    layout(
      title  = "Top 15 citizenship groups (descending)",
      xaxis  = list(title = "Examinees", showgrid = TRUE),
      yaxis  = list(title = "", autorange = "reversed"),
      height = max(420, nrow(top15_pc) * 34),
      margin = list(l = 150)
    )
}
```

```{r p04_pc_top15_tbl, echo=FALSE}
if (pc_available) dt_export(top15_pc, "Top 15 Citizenship Groups by Count")
```

```{r p04_pc_stacked}
if (pc_available && "PercentileBin" %in% names(pc_base)) {
  top_grps_pc <- (pc_base %>%
    dplyr::count(CITIZENSHIP_FINAL) %>%
    dplyr::arrange(dplyr::desc(n)) %>%
    head(15))$CITIZENSHIP_FINAL
  sub_pc_stk <- pc_base %>% dplyr::filter(CITIZENSHIP_FINAL %in% top_grps_pc)
  if (nrow(sub_pc_stk) > 0) {
    res_pc_stk <- pct_cross(sub_pc_stk, "CITIZENSHIP_FINAL", "PercentileBin", BIN_ORDER)
    long_pc_stk <- res_pc_stk$pct %>%
      tidyr::pivot_longer(-CITIZENSHIP_FINAL, names_to = "Bin", values_to = "Pct") %>%
      dplyr::mutate(Bin = factor(Bin, BIN_ORDER))
    plot_ly(long_pc_stk,
            x      = ~CITIZENSHIP_FINAL,
            y      = ~Pct,
            color  = ~Bin,
            type   = "bar",
            colors = PAL_BIN) %>%
      layout(barmode = "stack",
             title   = "Bin composition by citizenship (top 15)",
             xaxis   = list(title = "Citizenship"),
             yaxis   = list(title = "Percent"),
             height  = 520)
  }
}
```

```{r p04_pc_stacked_tbl, echo=FALSE}
if (pc_available && "PercentileBin" %in% names(pc_base)) {
  top15g <- (pc_base %>% dplyr::count(CITIZENSHIP_FINAL) %>%
               dplyr::arrange(dplyr::desc(n)) %>% head(15))$CITIZENSHIP_FINAL
  sub15  <- pc_base %>% dplyr::filter(CITIZENSHIP_FINAL %in% top15g)
  if (nrow(sub15) > 0) {
    res15 <- pct_cross(sub15, "CITIZENSHIP_FINAL", "PercentileBin", BIN_ORDER)
    dt_export(res15$pct, "Bin % by Citizenship Top 15 (row %)")
  }
}
```

```{r p04_pc_heatmap_all}
if (pc_available && "PercentileBin" %in% names(pc_base)) {
  res_pc_hm <- pct_cross(pc_base, "CITIZENSHIP_FINAL", "PercentileBin", BIN_ORDER)
  mat_pc_hm <- as.matrix(res_pc_hm$pct[, BIN_ORDER])
  rownames(mat_pc_hm) <- res_pc_hm$pct[["CITIZENSHIP_FINAL"]]
  plot_ly(
    z             = mat_pc_hm,
    x             = BIN_ORDER,
    y             = rownames(mat_pc_hm),
    type          = "heatmap",
    colorscale    = "YlOrRd",
    text          = round(mat_pc_hm, 1),
    texttemplate  = "%{text}",
    showscale     = TRUE
  ) %>%
    layout(
      title  = "B1-B10 Heatmap by Citizenship (Row %)",
      xaxis  = list(title = "Bin"),
      yaxis  = list(title = ""),
      height = max(400, nrow(mat_pc_hm) * 28)
    )
}
```

```{r p04_pc_heatmap_all_tbl, echo=FALSE}
if (pc_available && "PercentileBin" %in% names(pc_base)) {
  res_all <- pct_cross(pc_base, "CITIZENSHIP_FINAL", "PercentileBin", BIN_ORDER)
  dt_export(res_all$pct, "Full Bin % by Citizenship (row %)")
}
```

```{r p04_pc_topbin}
if (pc_available && "PercentileBin" %in% names(pc_base)) {
  td_pc <- pc_base %>%
    dplyr::mutate(is_top = PercentileBin %in% c("B8","B9","B10")) %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::summarise(n = dplyr::n(),
                     top_n = sum(is_top, na.rm = TRUE),
                     .groups = "drop") %>%
    dplyr::filter(n >= 3) %>%
    dplyr::mutate(top_pct = round(top_n / n * 100, 1)) %>%
    dplyr::arrange(top_pct)
  if (nrow(td_pc) > 0) {
    plot_ly(td_pc,
            x            = ~top_pct,
            y            = ~CITIZENSHIP_FINAL,
            type         = "bar",
            orientation  = "h",
            marker       = list(color = PAL_BIN[["B10"]]),
            text         = ~paste0(top_pct, "%"),
            textposition = "outside") %>%
      layout(
        title  = "Top-bin share (B8-B10) by citizenship",
        xaxis  = list(title = "% in B8-B10"),
        yaxis  = list(title = ""),
        height = max(360, nrow(td_pc) * 30)
      )
  }
}
```

```{r p04_pc_topbin_tbl, echo=FALSE}
if (pc_available && "PercentileBin" %in% names(pc_base)) {
  td2 <- pc_base %>%
    dplyr::mutate(is_top = PercentileBin %in% c("B8","B9","B10"),
                  is_bot = PercentileBin %in% c("B1","B2","B3")) %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::summarise(n       = dplyr::n(),
                     top_n   = sum(is_top, na.rm = TRUE),
                     bot_n   = sum(is_bot, na.rm = TRUE),
                     .groups = "drop") %>%
    dplyr::filter(n >= 3) %>%
    dplyr::mutate(top_pct = round(top_n / n * 100, 1),
                  bot_pct = round(bot_n / n * 100, 1)) %>%
    dplyr::arrange(dplyr::desc(top_pct))
  dt_export(td2, "Top & Bottom Bin Share by Citizenship")
}
```

```{r p04_pc_box_pct}
if (pc_available && "NMS_PER_num" %in% names(pc_base)) {
  big_pc_pct <- pc_base %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::filter(dplyr::n() >= 5) %>%
    dplyr::ungroup()
  if (nrow(big_pc_pct) > 0) {
    plot_ly(big_pc_pct,
            x         = ~CITIZENSHIP_FINAL,
            y         = ~NMS_PER_num,
            type      = "box",
            boxpoints = FALSE) %>%
      layout(
        title  = "Percentile rank by citizenship (n>=5)",
        xaxis  = list(title = ""),
        yaxis  = list(title = "Percentile rank"),
        height = 430
      )
  } else {
    plot_ly() %>% layout(title = "No data (n<5 for all groups)")
  }
}
```

```{r p04_pc_box_pct_tbl, echo=FALSE}
if (pc_available && "NMS_PER_num" %in% names(pc_base)) {
  stats_pct <- pc_base %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::filter(dplyr::n() >= 5) %>%
    dplyr::summarise(n      = dplyr::n(),
                     median = round(median(NMS_PER_num, na.rm = TRUE), 2),
                     q25    = round(quantile(NMS_PER_num, .25, na.rm = TRUE), 2),
                     q75    = round(quantile(NMS_PER_num, .75, na.rm = TRUE), 2),
                     .groups = "drop") %>%
    dplyr::arrange(dplyr::desc(median))
  dt_export(stats_pct, "Percentile Rank Stats by Citizenship (n>=5)")
}
```

```{r p04_pc_box_raw}
if (pc_available && "TotalRawScoreTRUE" %in% names(pc_base)) {
  big_pc_raw <- pc_base %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::filter(dplyr::n() >= 5) %>%
    dplyr::ungroup()
  if (nrow(big_pc_raw) > 0) {
    plot_ly(big_pc_raw,
            x         = ~CITIZENSHIP_FINAL,
            y         = ~TotalRawScoreTRUE,
            type      = "box",
            boxpoints = FALSE) %>%
      layout(
        title  = "TRUE raw score by citizenship (n>=5)",
        xaxis  = list(title = ""),
        yaxis  = list(title = "Total raw score"),
        height = 430
      )
  } else {
    plot_ly() %>% layout(title = "No data (n<5 for all groups)")
  }
}
```

```{r p04_pc_box_raw_tbl, echo=FALSE}
if (pc_available && "TotalRawScoreTRUE" %in% names(pc_base)) {
  stats_raw <- pc_base %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::filter(dplyr::n() >= 5) %>%
    dplyr::summarise(n      = dplyr::n(),
                     median = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
                     q25    = round(quantile(TotalRawScoreTRUE, .25, na.rm = TRUE), 2),
                     q75    = round(quantile(TotalRawScoreTRUE, .75, na.rm = TRUE), 2),
                     .groups = "drop") %>%
    dplyr::arrange(dplyr::desc(median))
  dt_export(stats_raw, "TRUE Raw Score Stats by Citizenship (n>=5)")
}
```

```{r p04_pc_summary}
if (pc_available) {
  has_bin_pc <- "PercentileBin" %in% names(pc_base)
  s_pc_sum <- pc_base %>%
    dplyr::mutate(
      is_top = if (has_bin_pc) PercentileBin %in% c("B8","B9","B10") else FALSE,
      is_bot = if (has_bin_pc) PercentileBin %in% c("B1","B2","B3")  else FALSE
    ) %>%
    dplyr::group_by(CITIZENSHIP_FINAL) %>%
    dplyr::summarise(
      n_examinees            = dplyr::n(),
      median_percentile_rank = round(median(NMS_PER_num, na.rm = TRUE), 2),
      median_true_raw_score  = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
      top_bin_n              = as.integer(sum(is_top, na.rm = TRUE)),
      bottom_bin_n           = as.integer(sum(is_bot, na.rm = TRUE)),
      .groups                = "drop"
    ) %>%
    dplyr::mutate(
      top_bin_pct    = round(top_bin_n    / pmax(n_examinees, 1) * 100, 2),
      bottom_bin_pct = round(bottom_bin_n / pmax(n_examinees, 1) * 100, 2)
    ) %>%
    dplyr::arrange(dplyr::desc(n_examinees), dplyr::desc(median_percentile_rank))
  dt_export(s_pc_sum, "Summary by citizenship")
}
```

```{r p04_pc_uni}
if (pc_available && "UNI_TYPE" %in% names(pc_base)) {
  s_pc_uni <- pc_base %>%
    dplyr::group_by(CITIZENSHIP_FINAL, UNI_TYPE) %>%
    dplyr::summarise(
      n                      = dplyr::n(),
      median_percentile_rank = round(median(NMS_PER_num, na.rm = TRUE), 2),
      median_true_raw_score  = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
      .groups                = "drop"
    )
  dt_export(s_pc_uni, "By citizenship and university type")
}
```

```{r p04_pc_course}
if (pc_available && "CourseGroup" %in% names(pc_base)) {
  s_pc_course <- pc_base %>%
    dplyr::group_by(CITIZENSHIP_FINAL, CourseGroup) %>%
    dplyr::summarise(
      n                      = dplyr::n(),
      median_percentile_rank = round(median(NMS_PER_num, na.rm = TRUE), 2),
      median_true_raw_score  = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
      .groups                = "drop"
    )
  dt_export(s_pc_course, "By citizenship and course group")
}
```

```{r p04_pc_year}
if (pc_available && "Year" %in% names(pc_base)) {
  s_pc_yr <- pc_base %>%
    dplyr::filter(!is.na(Year)) %>%
    dplyr::group_by(CITIZENSHIP_FINAL, Year) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    tidyr::pivot_wider(names_from = Year, values_from = n, values_fill = 0L)
  dt_export(s_pc_yr, "Year distribution by citizenship")
}
```

```{r p04_pc_records}
if (pc_available) {
  rec_cols_pc <- intersect(
    c("CITIZENSHIP_FINAL","FOREIGNER_STATUS","name_based_assessment",
      "PERSON_KEY","APPNO_CLEAN","UNIVERSITY","UNI_LOCATION","UNI_TYPE",
      "Year","CourseGroup","PercentileBin","NMS_PER_num",
      "TotalRawScoreTRUE","PLE_STATUS_LABEL"),
    names(pc_base)
  )
  s_pc_rec <- pc_base[, rec_cols_pc, drop = FALSE] %>%
    dplyr::arrange(CITIZENSHIP_FINAL, dplyr::desc(Year), dplyr::desc(NMS_PER_num))
  dt_export(s_pc_rec, "Matched no-PLE-match records with citizenship")
}
```

---

#### Comparative: Foreigners vs Filipino Groups

<div style="background:#e8f4f8;border-left:4px solid #3c8dbc;padding:12px 16px;margin-bottom:18px;border-radius:2px;font-size:14px;">
<b>Four comparison groups:</b><br>
<b>Foreigners</b> — no-PLE-match observable-cohort examinees profiled as foreign (FOREIGNER_STATUS == "Verified Foreigner").<br>
<b>Filipinos @ Foreign School</b> — profiled Filipinos (CITIZENSHIP_FINAL == "Filipino") enrolled at a Foreign university.<br>
<b>Public School</b> — all observable-cohort best-record examinees from Public universities.<br>
<b>Private School</b> — all observable-cohort best-record examinees from Private universities.
</div>

```{r p04_cmp_summary}
if (!is.null(cmp_df) && nrow(cmp_df) > 0) {
  cmp_sum <- cmp_df %>%
    dplyr::group_by(cmp_group) %>%
    dplyr::summarise(
      n                  = dplyr::n(),
      median_pct_rank    = round(median(NMS_PER_num,        na.rm = TRUE), 2),
      median_raw         = round(median(TotalRawScoreTRUE,  na.rm = TRUE), 2),
      top_bin_pct        = round(mean(PercentileBin %in% c("B8","B9","B10"), na.rm = TRUE) * 100, 2),
      ple_pass_rate_pct  = round(mean(HAS_CONFIRMED_PLE,    na.rm = TRUE) * 100, 2),
      .groups            = "drop"
    )
  dt_export(cmp_sum, "Comparative Summary: 4 Groups")
} else {
  dt_export(tibble::tibble(Note = "Comparison data unavailable (PC lookup file not found)."),
            "Comparative Summary")
}
```

```{r p04_cmp_box_pct}
if (!is.null(cmp_df) && nrow(cmp_df) > 0) {
  plot_ly(cmp_df,
          x         = ~cmp_group,
          y         = ~NMS_PER_num,
          type      = "box",
          color     = ~cmp_group,
          boxpoints = FALSE) %>%
    layout(
      title      = "Percentile Rank by Comparison Group",
      xaxis      = list(title = ""),
      yaxis      = list(title = "Percentile rank"),
      showlegend = FALSE
    )
} else {
  plot_ly() %>% layout(title = "No comparison data available")
}
```

```{r p04_cmp_heatmap}
if (!is.null(cmp_df) && nrow(cmp_df) > 0 && "PercentileBin" %in% names(cmp_df)) {
  res_cmp_hm <- pct_cross(cmp_df, "cmp_group", "PercentileBin", BIN_ORDER)
  mat_cmp_hm <- as.matrix(res_cmp_hm$pct[, BIN_ORDER])
  rownames(mat_cmp_hm) <- as.character(res_cmp_hm$pct[["cmp_group"]])
  plot_ly(
    z             = mat_cmp_hm,
    x             = BIN_ORDER,
    y             = rownames(mat_cmp_hm),
    type          = "heatmap",
    colorscale    = "RdYlGn",
    text          = round(mat_cmp_hm, 1),
    texttemplate  = "%{text}",
    showscale     = TRUE
  ) %>%
    layout(
      title  = "Full Bin Heatmap by Comparison Group (Row %)",
      xaxis  = list(title = "Bin"),
      yaxis  = list(title = "")
    )
} else {
  plot_ly() %>% layout(title = "No comparison data available")
}
```

```{r p04_cmp_heatmap_tbl, echo=FALSE}
if (!is.null(cmp_df) && nrow(cmp_df) > 0 && "PercentileBin" %in% names(cmp_df)) {
  res_cmp2 <- pct_cross(cmp_df, "cmp_group", "PercentileBin", BIN_ORDER)
  dt_export(res_cmp2$pct, "Bin % by Comparison Group (row %)")
}
```

```{r p04_cmp_stacked}
if (!is.null(cmp_df) && nrow(cmp_df) > 0 && "PercentileBin" %in% names(cmp_df)) {
  res_cmp_stk <- pct_cross(cmp_df, "cmp_group", "PercentileBin", BIN_ORDER)
  long_cmp_stk <- res_cmp_stk$pct %>%
    tidyr::pivot_longer(-cmp_group, names_to = "Bin", values_to = "Pct") %>%
    dplyr::mutate(Bin = factor(Bin, BIN_ORDER))
  plot_ly(long_cmp_stk,
          x      = ~cmp_group,
          y      = ~Pct,
          color  = ~Bin,
          type   = "bar",
          colors = PAL_BIN) %>%
    layout(barmode = "stack",
           title   = "Bin Composition by Comparison Group",
           xaxis   = list(title = "Group"),
           yaxis   = list(title = "Percent"))
} else {
  plot_ly() %>% layout(title = "No comparison data available")
}
```

```{r p04_cmp_stacked_tbl, echo=FALSE}
if (!is.null(cmp_df) && nrow(cmp_df) > 0 && "PercentileBin" %in% names(cmp_df)) {
  res_cmp3 <- pct_cross(cmp_df, "cmp_group", "PercentileBin", BIN_ORDER)
  dt_export(res_cmp3$counts, "Bin Counts by Comparison Group")
}
```

```{r p04_cmp_topbot}
if (!is.null(cmp_df) && nrow(cmp_df) > 0 && "PercentileBin" %in% names(cmp_df)) {
  cmp_tb <- cmp_df %>%
    dplyr::group_by(cmp_group) %>%
    dplyr::summarise(
      top_pct = round(mean(PercentileBin %in% c("B8","B9","B10"), na.rm = TRUE) * 100, 2),
      bot_pct = round(mean(PercentileBin %in% c("B1","B2","B3"),  na.rm = TRUE) * 100, 2),
      .groups = "drop"
    ) %>%
    tidyr::pivot_longer(c(top_pct, bot_pct), names_to = "Category", values_to = "Pct") %>%
    dplyr::mutate(Category = dplyr::recode(Category,
                                           top_pct = "Top bins B8-B10",
                                           bot_pct = "Bottom bins B1-B3"))
  plot_ly(cmp_tb,
          x      = ~cmp_group,
          y      = ~Pct,
          color  = ~Category,
          type   = "bar",
          colors = c("Top bins B8-B10" = "#2e7d32", "Bottom bins B1-B3" = "#c62828")) %>%
    layout(barmode    = "group",
           title      = "Top vs Bottom Bin % by Comparison Group",
           xaxis      = list(title = "Group"),
           yaxis      = list(title = "Percent"))
} else {
  plot_ly() %>% layout(title = "No comparison data available")
}
```

```{r p04_cmp_topbot_tbl, echo=FALSE}
if (!is.null(cmp_df) && nrow(cmp_df) > 0 && "PercentileBin" %in% names(cmp_df)) {
  dt_export(cmp_tb, "Top vs Bottom Bin % by Comparison Group")
}
```

## By Course Group

```{r p04_heatmap_course}
# Bin % heatmap by CourseGroup (YlOrRd)
res_cg <- pct_cross(F_besttrend, "CourseGroup", "PercentileBin", BIN_ORDER)
mat_cg <- as.matrix(res_cg$pct[, BIN_ORDER])
rownames(mat_cg) <- res_cg$pct[["CourseGroup"]]
plot_ly(
  z             = mat_cg,
  x             = BIN_ORDER,
  y             = rownames(mat_cg),
  type          = "heatmap",
  colorscale    = "YlOrRd",
  text          = round(mat_cg, 1),
  texttemplate  = "%{text}",
  showscale     = TRUE
) %>%
  layout(
    title  = "Bin % by Course Group",
    xaxis  = list(title = "Bin"),
    yaxis  = list(title = "")
  )
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Row percentages by course group and percentile bin.</p>

---

```{r p04_top_course}
# Top-bin share bar by CourseGroup
res_tc <- pct_cross(F_besttrend, "CourseGroup", "PercentileBin", BIN_ORDER)
top_cg <- res_tc$pct %>%
  dplyr::mutate(Top_B8_B10 = B8 + B9 + B10) %>%
  dplyr::select(CourseGroup, Top_B8_B10) %>%
  dplyr::arrange(Top_B8_B10)
plot_ly(top_cg,
        y           = ~CourseGroup,
        x           = ~Top_B8_B10,
        type        = "bar",
        orientation = "h",
        marker      = list(color = PAL_COURSE[top_cg$CourseGroup])) %>%
  layout(
    title  = "Top Bin Share by Course Group",
    xaxis  = list(title = "Percent in B8-B10"),
    yaxis  = list(title = "")
  )
```

```{r p04_top_course_tbl, echo=FALSE}
dt_export(top_cg, "Top Bin Share (B8-B10) by Course Group")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Top-bin (B8-B10) share for each pre-medical course background.</p>

---

```{r p04_course_tbl}
# Full bin % table by CourseGroup
res_cg_tbl <- pct_cross(F_besttrend, "CourseGroup", "PercentileBin", BIN_ORDER)
dt_export(res_cg_tbl$pct, "Bin % by Course Group")
```

<p style="color:#666;font-size:12px;margin:-6px 0 12px 0;">Full bin percentage table by course group.</p>

---

# University Type Analysis

## Overview

```{r p05_type_loc}
# Institution type x location table
df_p05 <- F_uni %>% dplyr::filter(!is.na(UNI_LOCATION))
tbl_p05 <- df_p05 %>%
  dplyr::count(UNI_TYPE, UNI_LOCATION) %>%
  dplyr::mutate(`% of total` = round(n / sum(n) * 100, 2))
dt_export(tbl_p05, "Institution Type x Location")
```

```{r p05_matrix}
# Count matrix: UNI_TYPE rows x UNI_LOCATION cols with marginals
df_p05b <- F_uni %>% dplyr::filter(!is.na(UNI_LOCATION))
tbl_mat <- as.data.frame.matrix(addmargins(table(df_p05b$UNI_TYPE, df_p05b$UNI_LOCATION)))
tbl_mat$UNI_TYPE <- rownames(tbl_mat)
rownames(tbl_mat) <- NULL
dt_export(tbl_mat, "Count Matrix: UNI_TYPE x UNI_LOCATION")
```

## Location Mix

```{r p05_inst_heatmap}
# Heatmap: bin % by Institution x Location (Blues)
df_p05c <- F_uni %>%
  dplyr::filter(!is.na(UNI_LOCATION), !is.na(PercentileBin)) %>%
  dplyr::mutate(inst = paste0(UNI_TYPE, " (", UNI_LOCATION, ")"))
res_p05c <- pct_cross(df_p05c, "inst", "PercentileBin", BIN_ORDER)
mat_p05c <- as.matrix(res_p05c$pct[, BIN_ORDER])
rownames(mat_p05c) <- res_p05c$pct$inst
plot_ly(
  z             = mat_p05c,
  x             = BIN_ORDER,
  y             = rownames(mat_p05c),
  type          = "heatmap",
  colorscale    = "Blues",
  texttemplate  = "%{z:.1f}"
) %>%
  layout(
    title  = "Bin % by Institution x Location",
    xaxis  = list(title = "Bin"),
    yaxis  = list(title = "")
  )
```

```{r p05_inst_heatmap_tbl, echo=FALSE}
dt_export(res_p05c$pct, "Bin % by Institution x Location (row %)")
```

```{r p05_top_bin_bar}
# Top-bin (B8-B10) share bar: Institution x Location
df_p05d <- F_uni %>%
  dplyr::filter(!is.na(UNI_LOCATION), !is.na(PercentileBin)) %>%
  dplyr::mutate(inst = paste0(UNI_TYPE, " (", UNI_LOCATION, ")"))
res_p05d <- pct_cross(df_p05d, "inst", "PercentileBin", BIN_ORDER)
top_p05d <- res_p05d$pct %>%
  dplyr::mutate(Top = B8 + B9 + B10) %>%
  dplyr::select(inst, Top) %>%
  dplyr::arrange(Top)
plot_ly(top_p05d,
        y           = ~inst,
        x           = ~Top,
        type        = "bar",
        orientation = "h") %>%
  layout(
    title  = "B8-B10 Share by Institution x Location",
    xaxis  = list(title = "Percent in B8-B10"),
    yaxis  = list(title = "")
  )
```

```{r p05_top_bin_bar_tbl, echo=FALSE}
dt_export(top_p05d, "Top Bin (B8-B10) Share by Institution x Location")
```

```{r p05_stacked_bin}
# Stacked bin bar by UNI_TYPE
res_p05e  <- pct_cross(F_uni, "UNI_TYPE", "PercentileBin", BIN_ORDER)
long_p05e <- res_p05e$pct %>%
  tidyr::pivot_longer(-UNI_TYPE, names_to = "Bin", values_to = "Pct") %>%
  dplyr::mutate(Bin = factor(Bin, BIN_ORDER))
plot_ly(long_p05e,
        x      = ~UNI_TYPE,
        y      = ~Pct,
        color  = ~Bin,
        type   = "bar",
        colors = PAL_BIN) %>%
  layout(barmode = "stack",
         title   = "Bin Composition by University Type",
         yaxis   = list(title = "Percent"),
         xaxis   = list(title = ""))
```

```{r p05_stacked_bin_tbl, echo=FALSE}
dt_export(res_p05e$pct, "Bin Composition by University Type (%)")
```

## Foreign Examinees

```{r p05_foreign_metrics, results="asis"}
# Inline metric cards for foreign examinees
foreign_p05 <- F_uni %>% dplyr::filter(UNI_TYPE == "Foreign")
n_for  <- nrow(foreign_p05)
pct_for <- round(n_for / max(nrow(F_uni), 1) * 100, 2)
med_for <- round(median(foreign_p05$NMS_PER_num, na.rm = TRUE), 1)
top_for <- round(mean(foreign_p05$PercentileBin %in% c("B8","B9","B10"), na.rm = TRUE) * 100, 2)

cat(sprintf(
  '<div style="display:flex;gap:16px;flex-wrap:wrap;margin-bottom:20px;">
    <div style="background:#f3e5f5;border-radius:6px;padding:14px 22px;min-width:160px;">
      <div style="font-size:26px;font-weight:700;color:#7b1fa2;">%s</div>
      <div style="font-size:12px;color:#555;">Foreign examinees</div>
    </div>
    <div style="background:#ede7f6;border-radius:6px;padding:14px 22px;min-width:160px;">
      <div style="font-size:26px;font-weight:700;color:#4527a0;">%s%%</div>
      <div style="font-size:12px;color:#555;">%% of total</div>
    </div>
    <div style="background:#e8eaf6;border-radius:6px;padding:14px 22px;min-width:160px;">
      <div style="font-size:26px;font-weight:700;color:#283593;">%s</div>
      <div style="font-size:12px;color:#555;">Median percentile rank</div>
    </div>
    <div style="background:#e3f2fd;border-radius:6px;padding:14px 22px;min-width:160px;">
      <div style="font-size:26px;font-weight:700;color:#1565c0;">%s%%</div>
      <div style="font-size:12px;color:#555;">Top bin (B8-B10) %%</div>
    </div>
  </div>',
  format(n_for,  big.mark = ","),
  pct_for,
  med_for,
  top_for
))
```

```{r p05_med_other}
# Medical & Allied vs Other by University Type
df_p05f <- F_uni %>%
  dplyr::mutate(bucket = ifelse(CourseGroup == "Medical & Allied",
                                "Medical & Allied", "Other Courses"))
res_p05f  <- pct_cross(df_p05f, "UNI_TYPE", "bucket")
long_p05f <- res_p05f$pct %>%
  tidyr::pivot_longer(-UNI_TYPE, names_to = "Bucket", values_to = "Pct")
plot_ly(long_p05f,
        x      = ~UNI_TYPE,
        y      = ~Pct,
        color  = ~Bucket,
        type   = "bar",
        colors = c("Medical & Allied" = "#d62728", "Other Courses" = "#7f7f7f")) %>%
  layout(barmode = "stack",
         title   = "Medical & Allied vs Other Courses by University Type",
         xaxis   = list(title = "University Type"),
         yaxis   = list(title = "Percent"))
```

```{r p05_med_other_tbl, echo=FALSE}
dt_export(res_p05f$pct, "Medical & Allied vs Other Courses by University Type (%)")
```

## Listings

```{r p05_tbl_public}
# Public university listing
make_uni_listing <- function(df, type_name) {
  sub <- df %>% dplyr::filter(UNI_TYPE == type_name)
  if (nrow(sub) == 0 || !"UNIVERSITY" %in% names(sub)) return(tibble::tibble())
  sub %>%
    dplyr::group_by(UNIVERSITY, UNI_LOCATION) %>%
    dplyr::summarise(Total_applicants = dplyr::n(), .groups = "drop") %>%
    dplyr::arrange(dplyr::desc(Total_applicants))
}
dt_export(make_uni_listing(F_uni, "Public"), "Public Universities")
```

```{r p05_tbl_private}
dt_export(make_uni_listing(F_uni, "Private"), "Private Universities")
```

```{r p05_tbl_foreign}
dt_export(make_uni_listing(F_uni, "Foreign"), "Foreign Universities")
```

---

# Flow & Pathways

## University to Bin

```{r p06_sankey_uni}
# Sankey: University Type -> Percentile Bin
flow_uni <- make_flow(F_uni, "UNI_TYPE", "PercentileBin",
                      c("Public","Private","Foreign"), BIN_ORDER)
make_sankey(flow_uni, "source", "target", "n", "University Type -> Bin")
```

```{r p06_flow_uni_tbl}
# Flow table: University -> Bin
flow_uni_tbl <- make_flow(F_uni, "UNI_TYPE", "PercentileBin")
dt_export(flow_uni_tbl, "Flow: University -> Bin")
```

## Course to Bin

```{r p06_sankey_course}
# Sankey: Course Group -> Percentile Bin
crs_order_p06 <- intersect(names(PAL_COURSE), unique(F_besttrend$CourseGroup))
flow_crs <- make_flow(F_besttrend, "CourseGroup", "PercentileBin", crs_order_p06, BIN_ORDER)
make_sankey(flow_crs, "source", "target", "n", "Course Group -> Bin")
```

```{r p06_flow_course_tbl}
# Flow table: Course -> Bin
flow_crs_tbl <- make_flow(F_besttrend, "CourseGroup", "PercentileBin")
dt_export(flow_crs_tbl, "Flow: Course -> Bin")
```

## Bin to PLE

```{r p06_sankey_ple}
# Sankey: Bin -> PLE Status (observable cohort only)
flow_ple <- make_flow(F_bestobs, "PercentileBin", "PLE_STATUS_LABEL",
                      BIN_ORDER, PLE_ORDER)
make_sankey(flow_ple, "source", "target", "n", "Bin -> PLE Status (Observable Cohort)")
```

```{r p06_flow_ple_tbl}
# Flow table with pct_pass column
flow_ple_tbl <- make_flow(F_bestobs, "PercentileBin", "PLE_STATUS_LABEL")
flow_ple_tbl$pct_pass <- round(
  flow_ple_tbl$n / ave(flow_ple_tbl$n, flow_ple_tbl$source, FUN = sum) * 100, 2
)
dt_export(flow_ple_tbl, "Flow: Bin -> PLE Status")
```

## Top Pathways

```{r p06_top_uni_pathways}
# Top pathways: University -> B8-B10
top_p06_uni <- F_uni %>%
  dplyr::filter(PercentileBin %in% c("B8","B9","B10")) %>%
  dplyr::count(UNI_TYPE, PercentileBin) %>%
  dplyr::arrange(dplyr::desc(n)) %>%
  head(10)
dt_export(top_p06_uni, "Top Pathways: University -> B8-B10")
```

```{r p06_top_course_pathways}
# Top pathways: Course -> B8-B10
top_p06_crs <- F_besttrend %>%
  dplyr::filter(PercentileBin %in% c("B8","B9","B10")) %>%
  dplyr::count(CourseGroup, PercentileBin) %>%
  dplyr::arrange(dplyr::desc(n)) %>%
  head(10)
dt_export(top_p06_crs, "Top Pathways: Course -> B8-B10")
```

# PLE Alignment

<div class="alert alert-info">
**Restricted to the observable best-record cohort (NMAT year &le; 2014)** to avoid misclassifying later cohorts as non-passers before their licensure window.
</div>

<div class="method-note">
**Score Profile tab:** Median and quartile scores split by PLE status. Mann-Whitney U tests each score dimension; effect size r &ge; 0.1 is small, &ge; 0.3 medium, &ge; 0.5 large. **Bin Profile tab:** Bar chart shows confirmed PLE pass rate per bin — the red dashed line marks 50%. **Background Links tab:** Top-bin survival by course group + PLE alignment by university type.
</div>

## Score Profile

```{r p07_score_profile}
score_cols_ple <- list(
  TotalRawScoreTRUE  = "Total Raw Score",
  PartIRawScoreTRUE  = "Part I",
  PartIIRawScoreTRUE = "Part II",
  NMS_PER_num        = "Percentile Rank",
  NMS_GPS            = "GPS Standard Score"
)
df_p07 <- F_bestobs
cols_p07 <- intersect(names(score_cols_ple), names(df_p07))
out_p07 <- df_p07 %>%
  filter(!is.na(PLE_STATUS_LABEL)) %>%
  group_by(PLE_STATUS_LABEL) %>%
  summarise(
    across(
      all_of(cols_p07),
      list(
        n      = ~ sum(!is.na(.)),
        median = ~ round(median(.x, na.rm = TRUE), 2),
        mean   = ~ round(mean(.x, na.rm = TRUE), 2),
        q25    = ~ round(quantile(.x, .25, na.rm = TRUE), 2),
        q75    = ~ round(quantile(.x, .75, na.rm = TRUE), 2)
      ),
      .names = "{.col}__{.fn}"
    ),
    .groups = "drop"
  )
dt_export(out_p07, "Score Profile by PLE Status")
```

```{r p07_boxplot}
df_p07_box <- F_bestobs %>%
  filter(!is.na(PLE_STATUS_LABEL), !is.na(TotalRawScoreTRUE))
plot_ly(
  df_p07_box,
  x     = ~PLE_STATUS_LABEL,
  y     = ~TotalRawScoreTRUE,
  type  = "box",
  color = ~PLE_STATUS_LABEL,
  colors = PAL_PLE,
  boxpoints = FALSE
) %>%
  layout(
    title      = "Total Raw Score by PLE Status",
    showlegend = FALSE,
    xaxis      = list(title = ""),
    yaxis      = list(title = "Total Raw Score")
  )
```

```{r p07_boxplot_tbl, echo=FALSE}
tbl_ple_scores <- F_bestobs %>%
  dplyr::filter(!is.na(PLE_STATUS_LABEL)) %>%
  dplyr::group_by(PLE_STATUS_LABEL) %>%
  dplyr::summarise(n            = dplyr::n(),
    median_raw   = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
    q25_raw      = round(quantile(TotalRawScoreTRUE, .25, na.rm = TRUE), 2),
    q75_raw      = round(quantile(TotalRawScoreTRUE, .75, na.rm = TRUE), 2),
    median_pct   = round(median(NMS_PER_num, na.rm = TRUE), 2),
    q25_pct      = round(quantile(NMS_PER_num, .25, na.rm = TRUE), 2),
    q75_pct      = round(quantile(NMS_PER_num, .75, na.rm = TRUE), 2),
    .groups      = "drop")
dt_export(tbl_ple_scores, "Score Statistics by PLE Status")
```

```{r p07_mw}
mw_col_map <- list(
  TotalRawScoreTRUE  = "Total Raw Score",
  PartIRawScoreTRUE  = "Part I",
  PartIIRawScoreTRUE = "Part II",
  NMS_PER_num        = "Percentile Rank",
  NMS_GPS            = "GPS Standard Score"
)
dt_export(mw_ple(F_bestobs, mw_col_map), "Mann-Whitney: PLE Status Comparison")
```

## Bin Profile

```{r p07_bin_stacked}
res_p07_bin <- pct_cross(F_bestobs, "PLE_STATUS_LABEL", "PercentileBin", BIN_ORDER)
long_p07_bin <- res_p07_bin$pct %>%
  pivot_longer(-PLE_STATUS_LABEL, names_to = "Bin", values_to = "Pct") %>%
  mutate(Bin = factor(Bin, BIN_ORDER))
plot_ly(
  long_p07_bin,
  x      = ~PLE_STATUS_LABEL,
  y      = ~Pct,
  color  = ~Bin,
  type   = "bar",
  colors = PAL_BIN
) %>%
  layout(
    barmode = "stack",
    title   = "Bin Distribution by PLE Status",
    xaxis   = list(title = ""),
    yaxis   = list(title = "Percent")
  )
```

```{r p07_bin_stacked_tbl, echo=FALSE}
dt_export(res_p07_bin$pct, "Bin Distribution by PLE Status (%)")
```

```{r p07_pass_bar}
df_p07_pass <- F_bestobs %>% filter(!is.na(PercentileBin))
res_p07_pass <- df_p07_pass %>%
  group_by(PercentileBin) %>%
  summarise(
    n        = n(),
    passers  = sum(PLE_STATUS_LABEL == "Confirmed PLE passer", na.rm = TRUE),
    pass_pct = round(passers / n * 100, 1),
    .groups  = "drop"
  ) %>%
  mutate(
    PercentileBin = factor(PercentileBin, BIN_ORDER),
    color         = PAL_BIN[as.character(PercentileBin)]
  )

plot_ly(
  res_p07_pass,
  x    = ~PercentileBin,
  y    = ~pass_pct,
  type = "bar",
  marker      = list(color = res_p07_pass$color),
  text        = ~paste0(pass_pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title      = "Confirmed PLE Pass Rate by NMAT Bin (Observable Cohort)",
    xaxis      = list(
      title          = "Percentile Bin",
      categoryorder  = "array",
      categoryarray  = BIN_ORDER
    ),
    yaxis      = list(title = "Confirmed PLE Pass Rate (%)", range = c(0, 105)),
    showlegend = FALSE,
    shapes = list(list(
      type = "line",
      x0   = -0.5, x1 = length(BIN_ORDER) - 0.5,
      y0   = 50,   y1 = 50,
      xref = "x",  yref = "y",
      line = list(color = "red", dash = "dash", width = 1.5)
    )),
    annotations = list(list(
      x        = length(BIN_ORDER) - 0.5,
      y        = 50,
      text     = "50%",
      showarrow = FALSE,
      xanchor  = "left",
      font     = list(color = "red", size = 11)
    ))
  )
```

```{r p07_pass_bar_tbl, echo=FALSE}
dt_export(res_p07_pass %>% dplyr::select(-color),
          "Confirmed PLE Pass Rate by NMAT Bin")
```

```{r p07_bin_tbl}
res_p07_bin_tbl <- pct_cross(F_bestobs, "PercentileBin", "PLE_STATUS_LABEL", PLE_ORDER)
dt_export(res_p07_bin_tbl$pct, "PLE Status % within Each Bin")
```

## Background Links

```{r p07_survival_tbl}
df_p07_surv <- F_besttrend %>% filter(!is.na(CourseGroup), !is.na(PercentileBin))
surv_p07 <- df_p07_surv %>%
  group_by(CourseGroup) %>%
  summarise(
    total   = n(),
    top_n   = sum(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE),
    top_pct = round(top_n / total * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(desc(top_pct))
dt_export(surv_p07, "Top-Bin Survival by Course Group")
```

```{r p07_survival_bar}
df_p07_sbar <- F_besttrend %>% filter(!is.na(CourseGroup), !is.na(PercentileBin))
s_p07_sbar <- df_p07_sbar %>%
  group_by(CourseGroup) %>%
  summarise(
    top_pct = round(mean(PercentileBin %in% c("B8", "B9", "B10"), na.rm = TRUE) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(top_pct)

plot_ly(
  s_p07_sbar,
  y           = ~CourseGroup,
  x           = ~top_pct,
  type        = "bar",
  orientation = "h",
  marker      = list(color = PAL_COURSE[s_p07_sbar$CourseGroup])
) %>%
  layout(
    title  = "B8–B10 Survival by Course Group",
    xaxis  = list(title = "% in B8–B10"),
    yaxis  = list(title = "")
  )
```

```{r p07_uni_ple}
tbl_p07_uni <- F_uniobs %>%
  group_by(UNI_TYPE) %>%
  summarise(
    n_observable = n(),
    confirmed    = sum(HAS_CONFIRMED_PLE, na.rm = TRUE),
    no_match     = n() - sum(HAS_CONFIRMED_PLE, na.rm = TRUE),
    pct          = round(mean(HAS_CONFIRMED_PLE, na.rm = TRUE) * 100, 2),
    .groups      = "drop"
  )
dt_export(tbl_p07_uni, "PLE Alignment by University Type")
```

## Policy Tables

```{r p07_pol_year}
dt_export(policy_tbls$by_year, "PLE Alignment by Year")
```

```{r p07_pol_course}
dt_export(policy_tbls$by_course, "PLE Alignment by Course")
```

```{r p07_pol_uni}
dt_export(policy_tbls$by_uni, "PLE Alignment by University Type")
```

# Repeat Takers

## Overview

```{r p08_attempt_bar}
ct_p08 <- repeat_data$attempt %>%
  count(n_attempts) %>%
  mutate(pct = round(n / sum(n) * 100, 2))

plot_ly(
  ct_p08,
  x    = ~n_attempts,
  y    = ~n,
  type = "bar",
  text = ~paste0(pct, "%"),
  textposition = "outside"
) %>%
  layout(
    title  = "NMAT Attempt Count Distribution",
    xaxis  = list(title = "Attempts"),
    yaxis  = list(title = "Count")
  )
```

```{r p08_attempt_tbl}
ct_p08_tbl <- repeat_data$attempt %>%
  count(n_attempts) %>%
  mutate(Percent = round(n / sum(n) * 100, 2))
names(ct_p08_tbl)[1:2] <- c("Attempts", "Count")
dt_export(ct_p08_tbl, "Attempt Count Distribution")
```

```{r p08_summary}
rd_p08 <- repeat_data
tbl_p08_sum <- tibble(
  Indicator = c(
    "Repeat-taker persons",
    "Analytic repeat takers",
    "Improved percentile (%)",
    "Improved raw score (%)",
    "Median percentile change",
    "Median raw score change",
    "Mean percentile change"
  ),
  Value = c(
    sum(rd_p08$attempt$n_attempts > 1),
    nrow(rd_p08$rep_df),
    round(mean(rd_p08$rep_df$pct_change > 0, na.rm = TRUE) * 100, 2),
    round(mean(rd_p08$rep_df$raw_change > 0, na.rm = TRUE) * 100, 2),
    round(median(rd_p08$rep_df$pct_change, na.rm = TRUE), 2),
    round(median(rd_p08$rep_df$raw_change, na.rm = TRUE), 2),
    round(mean(rd_p08$rep_df$pct_change, na.rm = TRUE), 2)
  )
)
dt_export(tbl_p08_sum, "Repeat-Taker Summary")
```

## Score Change

```{r p08_box_change}
rd_p08_long <- repeat_data$rep_df %>%
  select(pct_change, raw_change) %>%
  pivot_longer(everything(), names_to = "Measure", values_to = "Change") %>%
  mutate(
    Measure = recode(
      Measure,
      pct_change = "Percentile change",
      raw_change = "Raw score change"
    )
  )

plot_ly(
  rd_p08_long,
  x         = ~Measure,
  y         = ~Change,
  type      = "box",
  color     = ~Measure,
  boxpoints = FALSE
) %>%
  layout(
    title      = "Change from First to Last Attempt",
    showlegend = FALSE,
    yaxis      = list(title = "Change (last – first)"),
    shapes     = list(list(
      type = "line",
      x0   = -0.5, x1 = 1.5,
      y0   = 0,    y1 = 0,
      xref = "x",  yref = "y",
      line = list(color = "red", dash = "dash", width = 1.5)
    ))
  )
```

```{r p08_change_stats_tbl, echo=FALSE}
tbl_rpt_change <- repeat_data$rep_df %>%
  dplyr::summarise(
    n_repeat          = dplyr::n(),
    improved_pct_n    = sum(pct_change > 0, na.rm = TRUE),
    improved_pct_pct  = round(mean(pct_change > 0, na.rm = TRUE) * 100, 2),
    median_pct_change = round(median(pct_change, na.rm = TRUE), 2),
    mean_pct_change   = round(mean(pct_change, na.rm = TRUE), 2),
    sd_pct_change     = round(sd(pct_change, na.rm = TRUE), 2),
    improved_raw_n    = sum(raw_change > 0, na.rm = TRUE),
    improved_raw_pct  = round(mean(raw_change > 0, na.rm = TRUE) * 100, 2),
    median_raw_change = round(median(raw_change, na.rm = TRUE), 2),
    mean_raw_change   = round(mean(raw_change, na.rm = TRUE), 2),
    sd_raw_change     = round(sd(raw_change, na.rm = TRUE), 2)
  )
dt_export(tbl_rpt_change, "Repeat-Taker Score Change Summary Statistics")
```

```{r p08_scatter}
rd_p08_sc <- repeat_data$rep_df %>%
  filter(!is.na(first_pct), !is.na(last_pct))
lo_p08 <- min(rd_p08_sc$first_pct, rd_p08_sc$last_pct, na.rm = TRUE)
hi_p08 <- max(rd_p08_sc$first_pct, rd_p08_sc$last_pct, na.rm = TRUE)

plot_ly(
  rd_p08_sc,
  x      = ~first_pct,
  y      = ~last_pct,
  color  = ~n_attempts,
  type   = "scatter",
  mode   = "markers",
  marker = list(size = 5, opacity = .5),
  text   = ~paste0("Attempts: ", n_attempts),
  hovertemplate = "%{text}<br>First: %{x}<br>Last: %{y}<extra></extra>"
) %>%
  add_segments(
    x       = lo_p08, xend = hi_p08,
    y       = lo_p08, yend = hi_p08,
    inherit = FALSE,
    line    = list(color = "red", dash = "dash", width = 1.5),
    name    = "No change",
    showlegend = TRUE
  ) %>%
  layout(
    title  = "First vs Last Percentile — Repeat Takers",
    xaxis  = list(title = "First-attempt percentile"),
    yaxis  = list(title = "Last-attempt percentile")
  )
```

```{r p08_scatter_tbl, echo=FALSE}
dt_export(
  rd_p08_sc %>%
    dplyr::select(PERSON_KEY, n_attempts, first_year, last_year,
                  first_pct, last_pct, pct_change, first_raw, last_raw, raw_change) %>%
    dplyr::arrange(dplyr::desc(abs(pct_change))),
  "Repeat-Taker First vs Last Attempt - Full Detail"
)
```

## Detail Records

```{r p08_detail}
dt_export(
  repeat_data$rep_df %>% arrange(desc(pct_change)),
  "Repeat-Taker Detail"
)
```

# Subtest Profiles

## By University Type

```{r p09_heatmap_uni}
subtest_heatmap(
  subtest_means(
    F_uni %>% filter(UNI_TYPE %in% c("Public", "Private", "Foreign")),
    "UNI_TYPE"
  ),
  "Mean Std Subtest by University Type"
)
```

```{r p09_std_uni}
dt_export(
  subtest_means(F_uni, "UNI_TYPE"),
  "Std Subtest Means by University Type"
)
```

```{r p09_raw_uni}
dt_export(
  subtest_means(F_uni, "UNI_TYPE", use_std = FALSE),
  "Raw Subtest Means by University Type"
)
```

## By Course Group

```{r p09_heatmap_course}
subtest_heatmap(
  subtest_means(F_besttrend, "CourseGroup"),
  "Mean Std Subtest by Course Group"
)
```

```{r p09_std_course}
dt_export(
  subtest_means(F_besttrend, "CourseGroup"),
  "Std Subtest Means by Course Group"
)
```

```{r p09_raw_course}
dt_export(
  subtest_means(F_besttrend, "CourseGroup", use_std = FALSE),
  "Raw Subtest Means by Course Group"
)
```

## Radar Charts

```{r p09_radar_uni}
radar_plot(
  subtest_means(
    F_uni %>% filter(UNI_TYPE %in% c("Public", "Private", "Foreign")),
    "UNI_TYPE"
  ),
  "Radar: University Type"
)
```

```{r p09_radar_course}
radar_plot(
  subtest_means(F_besttrend, "CourseGroup"),
  "Radar: Course Group"
)
```

# Year Gap & Gender

## PLE Year Gap

```{r p10_gap_metrics, echo=FALSE, results="asis"}
gap_n   <- nrow(gap_df)
gap_med <- round(median(gap_df$PLE_YEAR_GAP, na.rm = TRUE), 1)
gap_q25 <- round(quantile(gap_df$PLE_YEAR_GAP, 0.25, na.rm = TRUE), 1)
gap_q75 <- round(quantile(gap_df$PLE_YEAR_GAP, 0.75, na.rm = TRUE), 1)

cat(sprintf(
  '<div style="display:flex;gap:16px;flex-wrap:wrap;margin-bottom:18px;">
    <div style="background:#e8f5e9;border-left:4px solid #2e7d32;padding:12px 20px;border-radius:4px;min-width:160px;">
      <div style="font-size:24px;font-weight:700;color:#2e7d32;">%s</div>
      <div style="font-size:13px;color:#555;">Confirmed passers with gap data</div>
    </div>
    <div style="background:#e8f5e9;border-left:4px solid #2e7d32;padding:12px 20px;border-radius:4px;min-width:160px;">
      <div style="font-size:24px;font-weight:700;color:#2e7d32;">%s</div>
      <div style="font-size:13px;color:#555;">Median year gap</div>
    </div>
    <div style="background:#e3f2fd;border-left:4px solid #1565c0;padding:12px 20px;border-radius:4px;min-width:160px;">
      <div style="font-size:24px;font-weight:700;color:#1565c0;">%s</div>
      <div style="font-size:13px;color:#555;">Q1 year gap</div>
    </div>
    <div style="background:#e3f2fd;border-left:4px solid #1565c0;padding:12px 20px;border-radius:4px;min-width:160px;">
      <div style="font-size:24px;font-weight:700;color:#1565c0;">%s</div>
      <div style="font-size:13px;color:#555;">Q3 year gap</div>
    </div>
  </div>',
  format(gap_n, big.mark = ","), gap_med, gap_q25, gap_q75
))
```

```{r p10_gap_hist}
plot_ly(gap_df, x = ~PLE_YEAR_GAP, type = "histogram", nbinsx = 20,
        marker = list(color = "#1f77b4")) %>%
  layout(title = "Distribution of NMAT-to-PLE Year Gap",
         xaxis = list(title = "Year gap"),
         yaxis = list(title = "Count"))
```

```{r p10_gap_hist_tbl, echo=FALSE}
tbl_gap_dist <- gap_df %>%
  dplyr::filter(!is.na(PLE_YEAR_GAP)) %>%
  dplyr::count(PLE_YEAR_GAP) %>%
  dplyr::mutate(pct = round(n / sum(n) * 100, 2)) %>%
  dplyr::arrange(PLE_YEAR_GAP)
dt_export(tbl_gap_dist, "NMAT-to-PLE Year Gap Distribution")
```

```{r p10_gap_box}
df_gap_box <- gap_df %>% filter(!is.na(CourseGroup))
plot_ly(df_gap_box, x = ~CourseGroup, y = ~PLE_YEAR_GAP, type = "box",
        color = ~CourseGroup, colors = PAL_COURSE, boxpoints = FALSE) %>%
  layout(title = "PLE Year Gap by Course Group", showlegend = FALSE,
         xaxis = list(title = "Course Group"),
         yaxis = list(title = "Year Gap"))
```

```{r p10_gap_tbl}
tbl_gap <- gap_df %>%
  group_by(CourseGroup) %>%
  summarise(
    confirmed_passers = n(),
    median_gap        = round(median(PLE_YEAR_GAP, na.rm = TRUE), 1),
    q25               = round(quantile(PLE_YEAR_GAP, 0.25, na.rm = TRUE), 1),
    q75               = round(quantile(PLE_YEAR_GAP, 0.75, na.rm = TRUE), 1),
    median_pct        = round(median(NMS_PER_num, na.rm = TRUE), 1),
    .groups           = "drop"
  ) %>%
  arrange(median_gap)
dt_export(tbl_gap, "Year-Gap Summary by Course Group")
```

## Gender Patterns

```{r p10_sex_summary}
tbl_sex <- sex_base %>%
  group_by(SEX_CLEAN) %>%
  summarise(
    n       = n_distinct(PERSON_KEY),
    med_raw = round(median(TotalRawScoreTRUE, na.rm = TRUE), 1),
    med_pct = round(median(NMS_PER_num, na.rm = TRUE), 1),
    med_gps = round(median(NMS_GPS, na.rm = TRUE), 1),
    .groups = "drop"
  )
dt_export(tbl_sex, "Score Summary by Sex")
```

```{r p10_sex_box}
plot_ly(sex_base, x = ~SEX_CLEAN, y = ~NMS_PER_num, type = "box",
        color = ~SEX_CLEAN,
        colors = c(Male = "#1f77b4", Female = "#e377c2"),
        boxpoints = FALSE) %>%
  layout(title = "Percentile Rank by Sex", showlegend = FALSE,
         xaxis = list(title = ""),
         yaxis = list(title = "Percentile Rank"))
```

```{r p10_sex_box_tbl, echo=FALSE}
tbl_sex_box <- sex_base %>%
  dplyr::group_by(SEX_CLEAN) %>%
  dplyr::summarise(n           = dplyr::n(),
    median_pct  = round(median(NMS_PER_num, na.rm = TRUE), 2),
    q25_pct     = round(quantile(NMS_PER_num, .25, na.rm = TRUE), 2),
    q75_pct     = round(quantile(NMS_PER_num, .75, na.rm = TRUE), 2),
    median_raw  = round(median(TotalRawScoreTRUE, na.rm = TRUE), 2),
    .groups     = "drop")
dt_export(tbl_sex_box, "Percentile Rank & Raw Score Statistics by Sex")
```

```{r p10_sex_year}
res_sex_yr  <- pct_cross(sex_base, "Year", "SEX_CLEAN", c("Male", "Female"))
long_sex_yr <- res_sex_yr$pct %>%
  pivot_longer(-Year, names_to = "Sex", values_to = "Pct")
plot_ly(long_sex_yr, x = ~Year, y = ~Pct, color = ~Sex, type = "bar",
        colors = c(Male = "#1f77b4", Female = "#e377c2")) %>%
  layout(barmode = "stack",
         title   = "Sex Composition by Year",
         yaxis   = list(title = "Percent"))
```

```{r p10_sex_yr_tbl, echo=FALSE}
tbl_sex_yr <- pct_cross(
  sex_base %>% dplyr::filter(!is.na(Year)),
  "Year", "SEX_CLEAN", c("Male", "Female")
)
dt_export(tbl_sex_yr$pct, "Sex Composition by NMAT Year (%)")
```

```{r p10_sex_ple_bar}
obs_sex      <- F_bestobs %>% filter(!is.na(SEX_CLEAN))
res_sex_ple  <- pct_cross(obs_sex, "SEX_CLEAN", "PLE_STATUS_LABEL", PLE_ORDER)
long_sex_ple <- res_sex_ple$pct %>%
  pivot_longer(-SEX_CLEAN, names_to = "Status", values_to = "Pct")
plot_ly(long_sex_ple, x = ~SEX_CLEAN, y = ~Pct, color = ~Status,
        type = "bar", colors = PAL_PLE) %>%
  layout(barmode = "stack",
         title   = "PLE Status by Sex (Observable Cohort)",
         xaxis   = list(title = ""),
         yaxis   = list(title = "Percent"))
```

```{r p10_sex_ple_tbl}
obs_sex2    <- F_bestobs %>% filter(!is.na(SEX_CLEAN))
res_sex_ple2 <- pct_cross(obs_sex2, "SEX_CLEAN", "PLE_STATUS_LABEL", PLE_ORDER)
dt_export(res_sex_ple2$pct, "PLE Status % by Sex")
```

# Statistical Tests

## Kruskal-Wallis

<div style="background:#fff8e1;border:1px solid #ffe082;padding:10px 14px;border-radius:4px;font-size:13px;margin-bottom:14px;">
<strong>Kruskal-Wallis H test:</strong> A non-parametric one-way ANOVA testing whether the distribution of each score metric differs significantly across NMAT exam years. This test makes no assumption about normality and is appropriate for the large, non-normal score distributions observed in this dataset. <strong>Eta-squared (&#951;&sup2;):</strong> effect size — 0.01 = small, 0.06 = medium, 0.14 = large. Values near zero indicate that year explains little of the variation in scores.
</div>

```{r p11_kw}
score_map_kw <- list(
  TotalRawScoreTRUE  = "Total raw score",
  PartIRawScoreTRUE  = "Part I raw score",
  PartIIRawScoreTRUE = "Part II raw score",
  NMS_PER_num        = "Percentile rank",
  NMS_GPS            = "GPS standard score"
)
dt_export(
  kw_table(F_besttrend, "Year", score_map_kw),
  "Kruskal-Wallis Across NMAT Years"
)
```

## Mann-Whitney

<div style="background:#fff8e1;border:1px solid #ffe082;padding:10px 14px;border-radius:4px;font-size:13px;margin-bottom:14px;">
<strong>Mann-Whitney U test:</strong> Compares score distributions between confirmed PLE passers and the no-confirmed-match group (observable cohort only, NMAT &le; 2014). This non-parametric test evaluates whether the two groups are drawn from the same distribution without assuming normality. <strong>Effect size r</strong> is computed as r = 1 &minus; 2U/(n&#8321;&times;n&#8322;): 0.1 = small, 0.3 = medium, 0.5 = large. Positive r means confirmed passers scored higher.
</div>

```{r p11_mw}
score_map_mw <- list(
  TotalRawScoreTRUE  = "Total raw score",
  PartIRawScoreTRUE  = "Part I raw score",
  PartIIRawScoreTRUE = "Part II raw score",
  NMS_PER_num        = "Percentile rank",
  NMS_GPS            = "GPS standard score"
)
dt_export(
  mw_ple(F_bestobs, score_map_mw),
  "Mann-Whitney by PLE Status (Observable Cohort)"
)
```

## Chi-Square

```{r p11_chi_obs}
df_chi    <- F_uni %>% filter(!is.na(UNI_TYPE), !is.na(PercentileBin))
ct_chi    <- as.data.frame.matrix(table(df_chi$UNI_TYPE, df_chi$PercentileBin))
for (b in setdiff(BIN_ORDER, names(ct_chi))) ct_chi[[b]] <- 0L
ct_chi$UNI_TYPE <- rownames(ct_chi)
rownames(ct_chi) <- NULL
avail_bins <- intersect(BIN_ORDER, names(ct_chi))
dt_export(ct_chi[, c("UNI_TYPE", avail_bins), drop = FALSE],
          "Observed Counts: UNI_TYPE x Bin")
```

```{r p11_chi_summary}
df_chi2 <- F_uni %>% filter(!is.na(UNI_TYPE), !is.na(PercentileBin))
ct_chi2 <- table(df_chi2$UNI_TYPE, df_chi2$PercentileBin)
tryCatch({
  ch   <- chisq.test(ct_chi2)
  n_ch <- sum(ct_chi2)
  V    <- sqrt(ch$statistic / (n_ch * min(nrow(ct_chi2) - 1, ncol(ct_chi2) - 1)))
  tbl_chi_sum <- tibble(
    chi2      = round(ch$statistic, 4),
    p_value   = ifelse(ch$p.value < 0.001, "<0.001", round(ch$p.value, 4)),
    df        = ch$parameter,
    n         = n_ch,
    cramers_v = round(V, 4)
  )
  dt_export(tbl_chi_sum, "Chi-Square Summary")
}, error = function(e) {
  dt_export(tibble(Error = conditionMessage(e)), "Chi-Square Summary")
})
```

```{r p11_chi_heatmap}
df_chi3   <- F_uni %>% filter(!is.na(UNI_TYPE), !is.na(PercentileBin))
ct_chi3   <- as.data.frame.matrix(table(df_chi3$UNI_TYPE, df_chi3$PercentileBin))
for (b in setdiff(BIN_ORDER, names(ct_chi3))) ct_chi3[[b]] <- 0L
avail3    <- intersect(BIN_ORDER, names(ct_chi3))
ct_sub3   <- ct_chi3[, avail3, drop = FALSE]
pct_chi3  <- round(sweep(ct_sub3, 1, pmax(rowSums(ct_sub3), 1), "/") * 100, 2)
plot_ly(z = as.matrix(pct_chi3), x = avail3, y = rownames(pct_chi3),
        type = "heatmap", colorscale = "YlGnBu",
        texttemplate = "%{z:.1f}") %>%
  layout(title  = "University Type x Bin (Row Percentages)",
         xaxis  = list(title = "Bin"),
         yaxis  = list(title = ""))
```

```{r p11_chi_heatmap_tbl, echo=FALSE}
pct_chi3_df <- as.data.frame(pct_chi3)
pct_chi3_df$UNI_TYPE <- rownames(pct_chi3_df)
rownames(pct_chi3_df) <- NULL
avail_b <- intersect(BIN_ORDER, names(pct_chi3_df))
dt_export(pct_chi3_df[, c("UNI_TYPE", avail_b), drop = FALSE],
          "University Type x Bin Row Percentages")
```

```{r p11_chi_exp}
df_chi4 <- F_uni %>% filter(!is.na(UNI_TYPE), !is.na(PercentileBin))
ct_chi4 <- table(df_chi4$UNI_TYPE, df_chi4$PercentileBin)
tryCatch({
  ch4   <- chisq.test(ct_chi4)
  exp4  <- as.data.frame(round(ch4$expected, 2))
  for (b in setdiff(BIN_ORDER, names(exp4))) exp4[[b]] <- NA_real_
  exp4$UNI_TYPE <- rownames(exp4)
  rownames(exp4) <- NULL
  avail4 <- intersect(BIN_ORDER, names(exp4))
  dt_export(exp4[, c("UNI_TYPE", avail4), drop = FALSE], "Expected Counts")
}, error = function(e) {
  dt_export(tibble(Error = conditionMessage(e)), "Expected Counts")
})
```

## Dunn Post-Hoc

```{r p11_dunn_check, echo=FALSE, results="asis"}
if (!HAS_DUNN) {
  cat('<div style="background:#fff3e0;border:1px solid #ffb300;padding:12px 16px;border-radius:4px;font-size:13px;">
  <strong>&#9888; dunn.test package not installed.</strong><br>
  To enable Dunn post-hoc pairwise comparisons, install the package and re-render this report:<br>
  <code>install.packages("dunn.test")</code>
  </div>')
} else {
  cat('<div style="background:#e3f2fd;border:1px solid #90caf9;padding:12px 16px;border-radius:4px;font-size:13px;margin-bottom:14px;">
  <strong>Dunn post-hoc test (Bonferroni adjustment):</strong><br>
  Pairwise comparisons of percentile rank distributions across all NMAT years (2006–2018).
  The table lists each year-pair, Z-statistic, and Bonferroni-adjusted p-value, sorted from most
  to least significant. The heatmap shows the full p-value matrix — darker cells (values near 0)
  indicate significantly different score distributions between those two years.
  </div>')
}
```

```{r p11_dunn_tbl, eval=HAS_DUNN}
df_dunn <- F_besttrend %>% filter(!is.na(Year), !is.na(NMS_PER_num))
res_dunn <- tryCatch({
  invisible(capture.output(
    r <- dunn.test::dunn.test(df_dunn$NMS_PER_num, as.character(df_dunn$Year),
                              method = "bonferroni", altp = TRUE)
  ))
  r
}, error = function(e) NULL)

if (!is.null(res_dunn)) {
  tbl_dunn <- tibble(
    Comparison  = res_dunn$comparisons,
    Z           = round(res_dunn$Z, 3),
    P_adjusted  = round(res_dunn$altP.adjusted, 5),
    Significant = ifelse(res_dunn$altP.adjusted < 0.05, "Yes", "No")
  ) %>% arrange(P_adjusted)
  dt_export(tbl_dunn, "Dunn Post-Hoc (Bonferroni-adjusted, sorted by p-value)")
} else {
  dt_export(tibble(Error = "Dunn test failed; check data."), "Dunn Post-Hoc")
}
```

```{r p11_dunn_heat, eval=HAS_DUNN}
df_dunn2 <- F_besttrend %>% filter(!is.na(Year), !is.na(NMS_PER_num))
yrs_d    <- sort(unique(df_dunn2$Year))
res_dunn2 <- tryCatch({
  invisible(capture.output(
    r <- dunn.test::dunn.test(df_dunn2$NMS_PER_num, as.character(df_dunn2$Year),
                              method = "bonferroni", altp = TRUE)
  ))
  r
}, error = function(e) NULL)

if (!is.null(res_dunn2)) {
  yrs_chr <- as.character(yrs_d)
  mat_d   <- matrix(NA_real_, nrow = length(yrs_chr), ncol = length(yrs_chr),
                    dimnames = list(yrs_chr, yrs_chr))
  diag(mat_d) <- 1
  comps_d <- strsplit(trimws(res_dunn2$comparisons), "\\s+-\\s+")
  for (i in seq_along(comps_d)) {
    a <- trimws(as.character(comps_d[[i]][1]))
    b <- trimws(as.character(comps_d[[i]][2]))
    if (a %in% rownames(mat_d) && b %in% colnames(mat_d)) {
      mat_d[a, b] <- res_dunn2$altP.adjusted[i]
      mat_d[b, a] <- res_dunn2$altP.adjusted[i]
    }
  }
  plot_ly(z = round(mat_d, 4), x = colnames(mat_d), y = rownames(mat_d),
          type = "heatmap", colorscale = "Viridis", reversescale = TRUE,
          texttemplate = "%{z:.3f}") %>%
    layout(title  = "Dunn Post-Hoc p-value Matrix (Percentile by Year)",
           xaxis  = list(title = "Year"),
           yaxis  = list(title = "Year"),
           height = 500)
} else {
  plot_ly() %>% layout(title = "Dunn test failed — check data.")
}
```

# Policy Tables & Export

<div style="background:#e8f4f8;border-left:4px solid #3c8dbc;padding:12px 16px;margin-bottom:18px;border-radius:2px;font-size:14px;">
These tables summarise PLE board exam alignment rates for the observable cohort (NMAT year &le; 2014). All tables include Copy, CSV, Excel, and Print export buttons. Since this is a static HTML report, use the export buttons in each table to download data.
</div>

```{r p12_year}
dt_export(policy_tbls$by_year, "Confirmed PLE Alignment by NMAT Year")
```

```{r p12_course}
dt_export(policy_tbls$by_course, "Confirmed PLE Alignment by Pre-Med Background")
```

```{r p12_uni}
dt_export(policy_tbls$by_uni, "Confirmed PLE Alignment by University Type")
```

```{r p12_survival}
dt_export(survival_tbl, "Top-Bin Survival by Course Group")
```

```{r p12_write_csvs, eval=FALSE, echo=TRUE}
# Run these lines interactively or in a separate script to export data as files.
# Requires the writexl or openxlsx package for Excel output.

write.csv(policy_tbls$by_year,   "ple_by_year.csv",   row.names = FALSE)
write.csv(policy_tbls$by_course, "ple_by_course.csv", row.names = FALSE)
write.csv(policy_tbls$by_uni,    "ple_by_uni.csv",    row.names = FALSE)
write.csv(survival_tbl,          "ple_survival.csv",  row.names = FALSE)

# Excel workbook with all four sheets (requires openxlsx):
# install.packages("openxlsx")
# wb <- openxlsx::createWorkbook()
# openxlsx::addWorksheet(wb, "By Year");    openxlsx::writeData(wb, 1, policy_tbls$by_year)
# openxlsx::addWorksheet(wb, "By Course");  openxlsx::writeData(wb, 2, policy_tbls$by_course)
# openxlsx::addWorksheet(wb, "By UniType"); openxlsx::writeData(wb, 3, policy_tbls$by_uni)
# openxlsx::addWorksheet(wb, "Survival");   openxlsx::writeData(wb, 4, survival_tbl)
# openxlsx::saveWorkbook(wb, "NMAT_Policy_Tables.xlsx", overwrite = TRUE)
```

---
*Report generated: `r format(Sys.time(), "%Y-%m-%d %H:%M")`. Data source: NMAT_Ultima.parquet. Observable cohort: NMAT year <= 2014 (n = `r nrow(F_bestobs)`). All analyses use best NMAT record per person.*
